---
title: "Coherence vignette"
author: "Erik Larsen"
date: "`r Sys.Date()`"
always_allow_html: true
output:
  html_document:
    toc: true
    fig_width: 10
    fig_height: 10
  pdf_document:
    fig_width: 10
    fig_height: 10
    latex_engine: xelatex
  md_document:
    variant: gfm
vignette: >
  %\VignetteIndexEntry{Coherence vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Overview

+ `hSAINT scores` are the **posterior probability** of a given protein being a 
*true interactor* with a target protein of interest

+ The `hierarchical Significance Analysis of INTeractome`, **hSAINT**, model-- a 
latent, 3-variable hierarachical Bayesian mixture model-- addresses some key 
limitations of the gold standard `SAINT` model
  
  + That model is a latent, 2-variable Poisson mixture model widely used in 
  "interactomics"
  
  + For more detail about the `hSAINT` model, see the [hSAINT README](../../python/README.md)

+ The goal of this vignette is to evaluate the **coherence** of clustering results
produced by multiple methods (`hierarchical clustering` and `HDBSCAN`) applied 
to IP‑MS/MS data

+ The data for the vignette was originally published in the `Journal of Cell Science`

# Introduction

**Coherence** is assessed along three independent axes:

1. **Statistical Coherence**  
   Whether proteins grouped into a cluster share similar `hSAINT` scores and
   whether the cluster exhibits a strong, interpretable `hSAINT` mass

2. **Contamination Coherence**  
   Whether the clean_fraction and contamination_rate of a cluster align with
   expectations for real biological signal (e.g., strong clusters may tolerate
   higher contamination if the clean subset is coherent)

3. **Semantic Coherence**  
   Whether the cluster yields enriched `GO term`s when using a contamination‑aware
   foreground definition

These three axes provide a principled way to determine whether a cluster 
represents real biological structure or is an artifact of the clustering method


```{r reset-shiny-mode, include=FALSE}
options(htmlwidgets.shinyMode = FALSE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.height = 12,
  dev = c("png", "pdf"),
  fig.path = "../../analysis/R/Coherence/"
)
```

## Attach Packages

```{r Attach Packages, message=F,warning=F}
# data/wrangling
library(tidyverse)
library(Endo)
library(DT)

  ## (hdbscan)
library(dbscan)

# GSEA/viz
library(clusterProfiler)
library(org.Hs.eg.db)
library(ReactomePA)
library(AnnotationDbi)
library(cowplot)

# progress bar
library(progressr)

# parallelization
library(future)
library(future.apply)

# network viz
library(visNetwork)

options(stringsAsFactors = FALSE)
```

## Data Inputs

+ Please see the other clustering vignettes on the classification methods used 
on proteins and clusters ([here](HDBSCAN.Rmd) and [here](HierarchicalClustering.Rmd))

```{r Data Load}
# Load precomputed objects from preprocessing
load("../../analysis/R/precompute_objects.rda")
load('../../analysis/R/hClust/hClust_protein_cluster_df.rda')
load('../../analysis/R/HDBSCAN/HDBSCAN_protein_cluster_df.rda')
load('../../analysis/R/hClust/HierarchicalSubClustCluster_GO_GSEA.rda')
load('../../analysis/R/hClust/HierarchicalCluster_GO_GSEA.rda')
load('../../analysis/R/HDBSCAN/HDBSCANSubClustCluster_GO_GSEA.rda')
load('../../analysis/R/HDBSCAN/HDBSCANCluster_GO_GSEA.rda')
```

```{r Create the Mean Shift Function to Characterize hSAINT scores}
mean_shift_valley <- function(gamma_vec) {

  # kernel density estimate on [0,1]
  d <- stats::density(gamma_vec, from = 0, to = 1)

  # tibble with density and neighbors for derivative and mode detection
  df <- tibble::tibble(
    x       = d$x,
    f       = d$y,
    f_lag   = dplyr::lag(d$y),      # f[i-1]
    f_lead  = dplyr::lead(d$y)      # f[i+1]
  )

  # discrete derivative of the density (forward difference)
  # slope[i] approximates f'(x[i])
  df <- df %>%
    dplyr::mutate(
      slope = f_lead - f
    )

  # indices where both neighbors exist (needed for local maxima)
  df <- df %>%
    dplyr::mutate(
      interior = !is.na(f_lag) & !is.na(f_lead)
    )

  # local maxima (aka modes) satisfy f > f_lag and f > f_lead
  df <- df %>%
    dplyr::mutate(
      is_mode = interior & f > f_lag & f > f_lead
    )

  # extract mode indices
  mode_idx <- which(df$is_mode)

  # assume two modes: left (noise) and right (signal)
  left_mode  <- mode_idx[1]
  right_mode <- mode_idx[length(mode_idx)]

  # mean-shift basin assignment:
  # each grid point moves uphill following the sign of the slope
  basin_target <- integer(nrow(df))

  for (i in seq_len(nrow(df))) {
    j <- i
    # move right while slope is positive
    while (j < nrow(df) && df$slope[j] > 0) {
      j <- j + 1
    }
    # move left while slope is negative
    while (j > 1 && df$slope[j - 1] < 0) {
      j <- j - 1
    }
    basin_target[i] <- j
  }

  # classify each point by which mode it's closer to
  left_dist  <- abs(basin_target - left_mode)
  right_dist <- abs(basin_target - right_mode)

  class_label <- ifelse(left_dist < right_dist, 1L, 2L)

  # valley boundary is the last index classified as left
  boundary_idx <- max(which(class_label == 1L))

  # valley location (single point)
  c <- df$x[boundary_idx]

  # expand valley asymmetrically using slope magnitude
  # threshold is relative to max slope magnitude
  eps <- 0.01 * max(abs(df$slope), na.rm = TRUE)

  left_idx <- boundary_idx
  while (left_idx > 1 && abs(df$slope[left_idx - 1]) < eps) {
    left_idx <- left_idx - 1
  }

  right_idx <- boundary_idx
  while (right_idx < nrow(df) && abs(df$slope[right_idx]) < eps) {
    right_idx <- right_idx + 1
  }

  c1 <- df$x[left_idx]
  c2 <- df$x[right_idx]

  c(c1, c2)
}
```

```{r Apply the Mean Shift}
hSAINT <- hSAINT |> 
  dplyr::mutate(
    hSAINT_class = {
      cutpoints <- mean_shift_valley(score) # find the indeces of the class modes bounding the valley between them
      dplyr::case_when(
        score < cutpoints[1] ~ 'Low hSAINT', # assign classes based on post. prob
        score <= cutpoints[2] ~ 'Medium hSAINT',
        score > cutpoints[2] ~ 'High hSAINT'
      )
    },
    contam_flag = lit_flag + CRAPome_flag,
    
    # assign contamination class based on frequency across shared studies and id in lit
    contam_class = dplyr::case_when(contam_flag == 0 ~ 
                                              'Low Contamination',
                                            contam_flag == 1 ~ 
                                              'Medium Contamination',
                                            contam_flag == 2 ~ 
                                              'High Contamination'),
    
    stat_class = paste(hSAINT_class,
                               contam_class, sep = ', ')
  ) |> # assign the orthogonal statistical model-based classes
  
  dplyr::rename(hSAINT_score = score) |> 
  dplyr::relocate(contam_flag, .after = hSAINT_score) |> 
  dplyr::select(-hSAINT_class, -contam_class) |> 
  dplyr::rename_with(.cols = c(!dplyr::contains('Protein')), ~paste0('protein_', .x))

head(hSAINT)
```


```{r Second precompute Data Load, eval = TRUE, echo = FALSE}
rm(IPMS_mat_log)
rm(IPMS_mat_Z)
```

```{r, results='asis', echo=FALSE}
fmt <- knitr::pandoc_to()

if (identical(fmt, "html")) {
  cat('# Cluster-Level Summaries {.tabset .tabset-pills .tabset-fade}\n')
} else if (identical(fmt, "markdown")) {
  cat('<details><summary><strong>Cluster-Level Summaries</strong></summary>\n')
} else {
  cat('# Cluster-Level Summaries\n')
}
```

## Cluster Assignments

Gather cluster-level summaries from both `hierarchical clustering` and 
`HDBSCAN clustering`

Summaries include:

  + cluster size
  + median `hSAINT` score
  + contamination rate
  + percentage of proteins not identified as contaminants
  + sum of `hSAINT` scores
  
These metrics form the foundation for coherence analysis

```{r Extract Cluster Info By Method}
cluster_assignments <- hClust_protein_cluster_df |> 
  dplyr::mutate(method = 'hClust', .before = Protein) |> 
  dplyr::full_join(
    HDBSCAN_protein_cluster_df |> 
      dplyr::mutate(method = 'HDBSCAN', .before = Protein)
  ) |> 
  dplyr::relocate(protein_stat_class, .after = cluster_class) |> 
  dplyr::group_by(method, Protein) |> 
  dplyr::arrange(Protein, method, protein_contam_flag, desc(protein_hSAINT_score))
```


```{r Cluster Summaries}
cluster_df <- cluster_assignments |> 
  dplyr::distinct(method, Protein, cluster, cluster_size, protein_hSAINT_score, protein_contam_flag) |> 
  dplyr::group_by(method, cluster) |> 
  dplyr::summarize(
    cluster_size = dplyr::n(),
    cluster_median_hSAINT_score = stats::median(protein_hSAINT_score, na.rm= T),
    cluster_mean_hSAINT_score = mean(protein_hSAINT_score,na.rm = T),
    cluster_contam_perc = (sum(protein_contam_flag)/cluster_size)*100,
    cluster_clean_fraction = (sum(protein_contam_flag == 0)/cluster_size)*100,
    cluster_hSAINT_score_mass = sum(protein_hSAINT_score, na.rm = T),
    cluster_clean_hSAINT_score_mass = sum(protein_hSAINT_score[protein_contam_flag == 0],na.rm = T )
  ) |> 
  dplyr::mutate(dplyr::across(c(is.double), ~round(.x, 2))) |> 
  dplyr::arrange(dplyr::desc(cluster_clean_fraction), dplyr::desc(cluster_median_hSAINT_score))
```

```{r Show Cluster Summaries, echo = FALSE}
DT::datatable(cluster_df, options = list(pageLength = 10))
```

## Plots

```{r cluster_median_hSAINT_score_vs_clean_fraction}
cluster_assignments |> 
  dplyr::group_by(method, cluster) |> 
  dplyr::distinct(cluster_size, cluster_median_hSAINT_score,
                  cluster_contam_perc) |> 
  dplyr::mutate(TMEM_Cluster = dplyr::case_when(cluster == 'noise' &
                                                  method == 'HDBSCAN' ~ 'TMEM',
                                                cluster == '12' &
                                                  method == 'hClust' ~ 'TMEM',
                                                TRUE ~ 'Not')) |> 
  ggplot2::ggplot(aes(x = cluster_median_hSAINT_score,
                      y = cluster_contam_perc, 
                      color = cluster, 
                      size = scale(cluster_size),
                      shape = TMEM_Cluster)) +
  ggplot2::geom_point() +
  ggplot2::labs(title = 'Cluster median hSAINT Score vs Cluster "Clean" Fraction',
                x = 'Cluster Median hSAINT Score',
                y = 'Cluster Contamination Percentage (%)',
                size = 'scale(cluster size)',
                shape = 'Is TMEM Cluster') +
  ggplot2::theme_bw() +
  ggplot2::theme(title = ggplot2::element_text(face = 'bold', size = 12),
                 axis.title = ggplot2::element_text(face = 'bold', size = 11),
                 axis.text = ggplot2::element_text(face = 'bold', size = 10),
                 legend.title = ggplot2::element_text(face = 'bold', size = 9),
                 legend.text = ggplot2::element_text(face = 'bold', size = 8),
                 strip.background = ggplot2::element_rect(fill = 'white', color = 'black'),
                 strip.text = ggplot2::element_text(face = 'bold', size = 8)) +
  ggplot2::facet_wrap(~method)
```

## Interpretation 

+ Bottom, right corners = "real biology"
+ Bottom, left corners = low hSAINT artifacts (keratins+)
+ Top = background; not candidates for GO enrichment
+ Middle = mixed clusters, containing some interesting proteins, but not coherent clusters

```{r, results='asis', echo=FALSE}
fmt <- knitr::pandoc_to()

if (identical(fmt, "html")) {
  cat('# Cluster-Level Coherence {.tabset .tabset-pills .tabset-fade}\n')
} else if (identical(fmt, "markdown")) {
  cat('<details><summary><strong>Cluster-Level Coherence</strong></summary>\n')
} else {
  cat('# Cluster-Level Coherence\n')
}
```

## hClust

+ `hClust` produces **28** clusters

+ `TMEM184B` is in **Cluster 12**

+ Background appears to be distributed across clusters with low 
`cluster_median_hSAINT_scores` with some common contaminants also appearing in 
**Cluster 25** and **Cluster 1**

```{r Show hClust Structure}
cluster_assignments |>
  dplyr::ungroup() |> 
  dplyr::filter(method == 'hClust') |> 
  dplyr::select(cluster, cluster_size, cluster_class, cluster_median_hSAINT_score) |> 
  dplyr::inner_join(cluster_assignments |>
                        dplyr::ungroup() |> 
                        dplyr::distinct(method, Protein, cluster) |> 
                        tidyr::pivot_wider(
                          id_cols = Protein,
                          names_from = method,
                          values_from = cluster
                          ) |> 
                        dplyr::count(hClust, HDBSCAN, name = 'n_overlap_with_HDBSCAN_clusters') |> 
                        dplyr::mutate(cluster = hClust) |> 
                        dplyr::rename(hClust_cluster = hClust, HDBSCAN_cluster = HDBSCAN)) |> 
  dplyr::select(-hClust_cluster)
```

## HDBSCAN

+ `hClust` produces **47** clusters

+ `TMEM184B` is in the **noise** cluster

+ Background appears to be consolidated within **Cluster 47**, though contaminants 
populate other high-scoring `hSAINT` clusters

```{r Show HDBSCAN Structure}
cluster_assignments |> 
  dplyr::ungroup() |> 
  dplyr::filter(method == 'HDBSCAN') |> 
  dplyr::select(cluster, cluster_size, cluster_class, cluster_median_hSAINT_score) |> 
  dplyr::inner_join(cluster_assignments |>
                        dplyr::ungroup() |> 
                        dplyr::distinct(method, Protein, cluster) |> 
                        tidyr::pivot_wider(
                          id_cols = Protein,
                          names_from = method,
                          values_from = cluster
                          ) |> 
                        dplyr::count(hClust, HDBSCAN, name = 'n_overlap_with_hClust_clusters') |> 
                        dplyr::mutate(cluster = HDBSCAN) |> 
                        dplyr::rename(HDBSCAN_cluster = HDBSCAN, hClust_cluster = hClust)) |> 
  dplyr::select(-HDBSCAN_cluster)
```

## Cross-Method Comparison

```{r Cross Method Comparison Mapping hClust Clusters to HDBSCAN}
# stable cluster projections:
  # how well do hClust clusters map to HDBSCAN clusters?
cluster_assignments |>
    dplyr::ungroup() |> 
    dplyr::distinct(method, Protein, cluster) |> 
    tidyr::pivot_wider(
      id_cols = Protein,
      names_from = method,
      values_from = cluster
      ) |> 
    dplyr::count(hClust, HDBSCAN, name = 'n_overlap') |> 
  dplyr::group_by(hClust) |> 
  dplyr::mutate(
    total = sum(n_overlap),
    frac = n_overlap / total
  ) |> 
  dplyr::slice_max(n_overlap, with_ties = F)
```


```{r Cross Method Comparison Mapping HDBSCAN Clusters to hClust}
  # how well do HDBSCAN clusters map to hClust clusters?
cluster_assignments |>
    dplyr::ungroup() |> 
    dplyr::distinct(method, Protein, cluster) |> 
    tidyr::pivot_wider(
      id_cols = Protein,
      names_from = method,
      values_from = cluster
      ) |> 
    dplyr::count(hClust, HDBSCAN, name = 'n_overlap') |> 
  dplyr::group_by(HDBSCAN) |> 
  dplyr::mutate(
    total = sum(n_overlap),
    frac = n_overlap / total
  ) |> 
  dplyr::slice_max(n_overlap, with_ties = F)
```

## Interpretation

Many `hClust` clusters map primarily to the `HDBSCAN` noise/background clusters 
(**"noise"**, **47**)

  + this is what `HDBSCAN` is designed to do
    + i.e. dense pockets as clusters, label everything else as noise
    
  + these are low-`hSAINT score`-signal `hClust` clusters and with contaminants
  
The **TMEM** `hClust` cluster (**Cluster 12**), which had no significantly enriched GO terms
and a very small `cluster median hSAINT score`, is classified in the 
`low-hSAINT, low-contamination` quadrant with no significantly enriched GO terms

These proteins map to the **noise Cluster** under `HDBSCAN`, a comparatively much larger (**2903** vs **161**) 
`high-hSAINT, high-contamination` background cluster

This mapping reflects geometric density behavior rather than biological similarity:

  + `TMEM` proteins are low-signal and sparse, and `HDBSCAN` assigns such points
  to the nearest dense region rather than forming a separate cluster

```{r}

```

```{r, results='asis', echo=FALSE}
fmt <- knitr::pandoc_to()

if (identical(fmt, "html")) {
  cat('# GO Enrichment {.tabset .tabset-pills .tabset-fade}\n')
} else if (identical(fmt, "markdown")) {
  cat('<details><summary><strong>GO Enrichment</strong></summary>\n')
} else {
  cat('# GO Enrichment\n')
}
```


```{r, results='asis', echo=FALSE}
fmt <- knitr::pandoc_to()

if (identical(fmt, "html")) {
  cat('# Semantic Coherence {.tabset .tabset-pills .tabset-fade}\n')
} else if (identical(fmt, "markdown")) {
  cat('<details><summary><strong>Semantic Coherence</strong></summary>\n')
} else {
  cat('# Semantic Coherence\n')
}
```
