---
title: "HDBSCAN vignette"
author: "Erik Larsen"
date: "`r Sys.Date()`"
always_allow_html: true
output:
  html_document:
    toc: true
    fig_width: 10
    fig_height: 10
  pdf_document:
    fig_width: 10
    fig_height: 10
    latex_engine: xelatex
  md_document:
    variant: gfm
vignette: >
  %\VignetteIndexEntry{HDBSCAN vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Overview

This markdown was developed to formally document an ad-hoc cluster analysis for 
TMEM184B immunoprecipitation tandem mass-spec experiments for 
**Dr. Martha Bhattacharya's lab** at the **University of Arizona** for the
endolysosomal paper under revision (2025-02-28) for the Journal of Cell Science

This vignette documents the `Spectral Clustering` of this IPMS data

```{r reset-shiny-mode, include=FALSE}
options(htmlwidgets.shinyMode = FALSE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.height = 12,
  dev = c("png", "pdf"),
  fig.path = "../../analysis/R/SpectralClustering/"
)
```

## Attach Packages

```{r Attach Packages, message=F,warning=F}
# data/wrangling
library(tidyverse)
library(Endo)
library(DT)

  ## (hdbscan)
library(dbscan)

# GSEA/viz
library(clusterProfiler)
library(org.Hs.eg.db)
library(ReactomePA)
library(AnnotationDbi)
library(cowplot)

# progress bar
library(progressr)

# parallelization
library(future)
library(future.apply)

# network viz
library(visNetwork)

options(stringsAsFactors = FALSE)
```

## Load the Pre-Computed Data

```{r Data Load}
# Load precomputed objects from preprocessing
load("../../analysis/R/precompute_objects.rda")
```

```{r Second Data Load, echo = FALSE}
rm(IPMS_mat_log)
rm(IPMS_mat_Z)
hSAINT <- hSAINT |> 
  dplyr::mutate(stat_class = dplyr::case_when(score > 0.5 & 
                                                lit_flag + CRAPome_flag <= 1 ~ 'High SAINT, Low Contamination',
                                              score < 0.5 &
                                                lit_flag + CRAPome_flag <= 1 ~ 'Low SAINT, Low Contamination',
                                              score > 0.5 &
                                                lit_flag + CRAPome_flag > 1 ~ 'High SAINT, High Contamination',
                                              score < 0.5 &
                                                lit_flag + CRAPome_flag > 1 ~ 'Low SAINT, High Contamination'))
```

