---
title: "HDBSCAN vignette"
author: "Erik Larsen"
date: "`r Sys.Date()`"
always_allow_html: true
output:
  html_document:
    toc: true
    fig_width: 10
    fig_height: 10
  pdf_document:
    fig_width: 10
    fig_height: 10
    latex_engine: xelatex
  md_document:
    variant: gfm
vignette: >
  %\VignetteIndexEntry{HDBSCAN vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Overview

This markdown was developed to formally document an ad-hoc cluster analysis for 
TMEM184B immunoprecipitation tandem mass-spec experiments for 
**Dr. Martha Bhattacharya's lab** at the **University of Arizona** for the
endolysosomal paper under revision (2025-02-28) for the Journal of Cell Science

This vignette documents the `HDBSCAN` of this IPMS data

```{r reset-shiny-mode, include=FALSE}
options(htmlwidgets.shinyMode = FALSE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.height = 12,
  dev = c("png", "pdf"),
  fig.path = "../../analysis/R/SpectralClustering/"
)
```

## Attach Packages

```{r Attach Packages, message=F,warning=F}
# data/wrangling
library(tidyverse)
library(Endo)
library(DT)

  ## (hdbscan)
library(dbscan)

# GSEA/viz
library(clusterProfiler)
library(org.Hs.eg.db)
library(ReactomePA)
library(AnnotationDbi)
library(cowplot)

# progress bar
library(progressr)

# parallelization
library(future)
library(future.apply)

# network viz
library(visNetwork)

options(stringsAsFactors = FALSE)
```

## Load the Pre-Computed Data

```{r Data Load}
# Load precomputed objects from preprocessing
load("../../analysis/R/precompute_objects.rda")
```

```{r Second Data Load, echo = FALSE}
rm(IPMS_mat_log)
rm(IPMS_mat_Z)
hSAINT <- hSAINT |> 
  dplyr::mutate(stat_class = dplyr::case_when(score > 0.5 & 
                                                lit_flag + CRAPome_flag <= 1 ~ 'High SAINT, Low Contamination',
                                              score < 0.5 &
                                                lit_flag + CRAPome_flag <= 1 ~ 'Low SAINT, Low Contamination',
                                              score > 0.5 &
                                                lit_flag + CRAPome_flag > 1 ~ 'High SAINT, High Contamination',
                                              score < 0.5 &
                                                lit_flag + CRAPome_flag > 1 ~ 'Low SAINT, High Contamination'))
```

```{r, results='asis', echo=FALSE}
fmt <- knitr::pandoc_to()

if (identical(fmt, "html")) {
  cat('# Principal Component Analysis {.tabset .tabset-pills .tabset-fade}\n')
} else if (identical(fmt, "markdown")) {
  cat('<details><summary><strong>Principal Component Analysis</strong></summary>\n')
} else {
  cat('# Principal Component Analysis\n')
}
```
Conduct **PCA** on the `IPMS counts dataframe` to extract the 2-D matrix required by 
`hdbscan`

```{r PCA on Counts}
pca <- stats::prcomp(IPMS_counts[,2:ncol(IPMS_counts)], scale = T)

X <- pca$x[ , 1:2, drop = FALSE]
```

```{r, results='asis', echo=FALSE}
fmt <- knitr::pandoc_to()

if (identical(fmt, "html")) {
  cat('# Clustering {.tabset .tabset-pills .tabset-fade}\n')
} else if (identical(fmt, "markdown")) {
  cat('<details><summary><strong>Clustering</strong></summary>\n')
} else {
  cat('# Clustering\n')
}
```

## Detect Clusters

Pass the two-dimensional object through `HDBSCAN` to obtain clustering results

+ set `minPts` (minimum number of points required to form a dense region) to 
default (**10**), which provides a stable balance between detecting 
meaningful dense regions and avoiding fragmentation

`HDBSCAN` will identify clusters from the 2-D object as regions of high point 
density within the embedding

```{r Scan the PCA results}
hdb <- dbscan::hdbscan(x = X, minPts = 10)
```

## Extract Clusters

```{r Extract Clusters}
cluster_labels <- hdb$cluster
```

## Tabulate Clusters

```{r Tabulate Clusters}
cluster_table <- table(cluster_labels)

cluster_labels |> 
  as.data.frame() |> 
  dplyr::mutate(cluster_labels = as.character(cluster_labels)) |> 
  dplyr::group_by(cluster_labels) |> 
  dplyr::summarize(n = dplyr::n()) |> 
  dplyr::arrange(desc(n)) |> 
  dplyr::mutate(cluster_labels = dplyr::case_when(cluster_labels == '0' ~ 'noise',
                                                  TRUE ~ cluster_labels)) |> 
  head()
```

```{r, results='asis', echo=FALSE}
fmt <- knitr::pandoc_to()

if (identical(fmt, "html")) {
  cat('# Cluster Visualization {.tabset .tabset-pills .tabset-fade}\n')
} else if (identical(fmt, "markdown")) {
  cat('<details><summary><strong>Cluster Visualization</strong></summary>\n')
} else {
  cat('# Cluster Visualization\n')
}
```

Plot the PCs

## PCA Plot

```{r PCA Plot}
pca_results_df <- tibble::tibble(
  PC1 = X[ , 1],
  PC2 = X[ , 2],
  cluster = cluster_labels |> 
                     as.data.frame() |> 
                     dplyr::mutate(cluster_labels = as.character(
                       cluster_labels
                     ),
                     cluster_labels = dplyr::case_when(cluster_labels == '0' ~ 
                                                         'noise',
                                                       TRUE ~ 
                                                         cluster_labels)) |> 
                     dplyr::pull()
)

cowplot::plot_grid(
  pca_results_df |> 
    ggplot2::ggplot(ggplot2::aes(x = PC1, y = PC2, group = cluster, color = cluster)) +
    ggplot2::geom_point(size = 2, alpha = 0.3) +
    ggplot2::labs(
      title = 'PCA of HDBSCAN Clusters'
      ) +
    ggplot2::theme_bw() +
    ggplot2::theme(title = ggplot2::element_text(face = 'bold', size = 12),
                   axis.title = ggplot2::element_text(face = 'bold', size = 10),
                   axis.text = ggplot2::element_text(face = 'bold', size = 9),
                   legend.position = 'bottom', 
                   legend.direction = 'horizontal',
                   legend.spacing.x = unit(0.5, 'cm'))
  ,
  pca_results_df |> 
    dplyr::filter(cluster != 'noise') |> 
    ggplot2::ggplot(ggplot2::aes(x = PC1, y = PC2, group = cluster, color = cluster)) +
    ggplot2::geom_point(size = 2, alpha = 0.3) +
    ggplot2::labs(
      title = 'PCA of HDBSCAN Clusters\nwithout "noise" cluster'
      ) +
    ggplot2::theme_bw() +
    ggplot2::theme(title = ggplot2::element_text(face = 'bold', size = 12),
                   axis.title = ggplot2::element_text(face = 'bold', size = 10),
                   axis.text = ggplot2::element_text(face = 'bold', size = 9),
                   legend.position = 'bottom', 
                   legend.direction = 'horizontal',
                   legend.spacing.x = unit(0.5, 'cm'))
    ,
  ncol = 2,
  align = 'h'
)
```

```{r, results='asis', echo=FALSE}
fmt <- knitr::pandoc_to()

if (identical(fmt, "html")) {
  cat('# Cluster Descriptions {.tabset .tabset-pills .tabset-fade}\n')
} else if (identical(fmt, "markdown")) {
  cat('<details><summary><strong>Cluster Descriptions</strong></summary>\n')
} else {
  cat('# Cluster Descriptions\n')
}
```


In order to assess this method and also to compare/validate the hierarchical SAINT
model, describing clusters based on the proteins in them is useful

Similarly localized proteins could be expected to have similar abundances to 
each other as well as the target (**TMEM184B**)

Similar functioning proteins could also be expected to have similar abundances 
to each other as well as the target

Describing these clusters attained by `HDBSCAN` via 
gene ontology gene set enrichment analysis is, therefore, useful and standard 
practice

Furthermore, we may be able to assemble a `protein-protein interaction network`
from the results

I outline this below-- the methods and results for acquiring these clusters, 
their constituent proteins, and how we make sense of them-- visualized as
`dotplots`, `heatmaps`, and a `network`. I include searchable tables printed last

## Extract Clusters

Extract gene names from the hierarchically-clustered clusters and convert to a 
dataframe

```{r Extract Gene Names from Clusters}
clusters_df <- pca_results_df |> 
  dplyr::select(cluster) |> 
  dplyr::mutate(Protein = IPMS_counts |> dplyr::pull(Protein), .before = 1)

head(clusters_df)
```

```{r, echo = FALSE}
rm(list = ls(pattern = '_mat'))
```

## Join Clusters to hSAINT

Include score summaries (mean and median posterior probability scores, ranks, and literature or CRAPome contamination rate)

```{r Summarize SAINT Scores and Contamination Rates}
cluster_scores <- clusters_df |>  
  dplyr::inner_join(
    hSAINT
  ) |> 
  dplyr::arrange(desc(score)) |> 
  dplyr::group_by(cluster) |> 
  dplyr::summarize(mean_score = mean(score),
                   median_score = stats::median(score),
                   score_mass = sum(score),
                   n = dplyr::n(),
                   lit_flag_perc = (sum(lit_flag)/n)*100,
                   CRAPome_flag_perc = (sum(CRAPome_flag)/n)*100,
                   contam_perc = lit_flag_perc + CRAPome_flag_perc) |> 
  dplyr::arrange(desc(mean_score), contam_perc) |> 
  dplyr::mutate(mean_rank = dplyr::row_number()) |> 
  dplyr::arrange(desc(median_score), contam_perc) |> 
  dplyr::mutate(median_rank = dplyr::row_number()) |> 
  dplyr::arrange(desc(score_mass), contam_perc) |> 
  dplyr::mutate(total_rank = dplyr::row_number()) |> 
  dplyr::arrange(mean_rank) |> 
  dplyr::relocate(c(lit_flag_perc, CRAPome_flag_perc), .after = total_rank)

head(cluster_scores)
```

Join the cluster scores to the cluster (per protein) dataframe and the `hSAINT` results for classification

```{r Join the Summary dataframe to the Protein Level dataframes}
protein_cluster_df <- clusters_df |> 
  dplyr::inner_join(
    hSAINT
  ) |> 
  dplyr::full_join(
    cluster_scores |> 
      dplyr::ungroup()
  ) |> 
  dplyr::mutate(
    cluster_SAINT_class = dplyr::case_when(mean_score > 0.5 ~ 'High SAINT',
                                           mean_score < 0.5 ~ 'Low SAINT'
                                            ),
    
    cluster_contams_class = dplyr::case_when(contam_perc > cluster_scores |> 
                                               dplyr::select(contam_perc) |> 
                                               dplyr::pull() |> 
                                               stats::median() ~
                                               'High Contamination',
                                             TRUE ~ 
                                               'Low Contamination')) |> 
  dplyr::mutate(cluster_class = paste0(cluster_SAINT_class,
                                        ", ",
                                        cluster_contams_class)) |> 
  dplyr::select(-cluster_SAINT_class,
                -cluster_contams_class) |> 
  dplyr::relocate(cluster_class, .after = cluster) |> 
  dplyr::arrange(desc(score), Protein)

head(protein_cluster_df)
```

## Cluster GO GSEA

```{r Downstream Analysis}
EntrezIDs <- AnnotationDbi::select(
  org.Hs.eg.db,
  keys = clusters_df |> dplyr::distinct(Protein) |> dplyr::pull(),
  columns = 'ENTREZID',
  keytype = 'SYMBOL'
) |> 
  dplyr::filter(!is.na(2))

EntrezMap <- stats::setNames(EntrezIDs$ENTREZID, EntrezIDs$SYMBOL)
```

Create enrichment functions

```{r Create Enrichment Functions}
background_genes <- EntrezIDs |> # use the EntrezIDs mapped from the modules_df
  dplyr::pull(ENTREZID)

get_GO <- function(genes, ontology, background_genes) {
  clusterProfiler::enrichGO(
    gene = genes,
    universe = background_genes,
    OrgDb = org.Hs.eg.db,
    keyType = 'ENTREZID',
    ont = ontology,
    pAdjustMethod = 'BH',
    qvalueCutoff = 0.05,
    readable = T
  )
}

get_kegg <- function(genes, EntrezMap) {
  entrez <- EntrezMap[genes]
  entrez <- entrez[!is.na(entrez)]
  if (length(entrez)==0) {
    return(NULL)
  }
  clusterProfiler::enrichKEGG(
    gene = entrez,
    organism = 'hsa',
    pAdjustMethod = 'BH',
    qvalueCutoff = 0.05
  )
}

get_reactome <- function(genes, EntrezMap) {
  entrez <- EntrezMap[genes]
  entrez <- entrez[!is.na(entrez)]
  if (length(entrez)==0) {
    return(NULL)
  }
  ReactomePA::enrichPathway(
    gene = entrez,
    organism = 'human',
    pAdjustMethod = 'BH',
    qvalueCutoff = 0.05,
    readable = T
  )
}
```

Run the functions on the dataset in parallel (multiple CPUs)

```{r Parallelize Enrichment, eval = FALSE}
future::plan(future::multisession, workers = 6)
progressr::handlers(progressr::handler_txtprogressbar())

cluster_ids <- sort(unique(clusters_df$cluster))

progressr::with_progress({
  p <- progressr::progressor(steps = length(cluster_ids))

  enrichment_results <- future.apply::future_lapply(
    cluster_ids,
    function(mod) {

      # tick the progress bar
      p(sprintf("cluster %s", mod))

      genes <- clusters_df |>
        dplyr::filter(cluster == mod) |>
        dplyr::inner_join(EntrezIDs |> dplyr::rename(Protein = 1)) |> 
        dplyr::distinct(ENTREZID) |>
        dplyr::pull(ENTREZID)

      list(
        cluster = mod,
        go_bp  = get_GO(genes, ontology = "BP", background_genes = background_genes),
        go_cc  = get_GO(genes, ontology = "CC", background_genes = background_genes),
        go_mf  = get_GO(genes, ontology = "MF", background_genes = background_genes)
        # add KEGG/Reactome later
      )
    }
  )
})

names(enrichment_results) <- cluster_ids
rm(p)
```

Gather enrichment results

```{r Consolidate Enrichment Results, eval = FALSE}
global_enrichment_results_df <- lapply(enrichment_results, function(x) {
  x %>%
    purrr::map_df(., .f = as.data.frame, .id = 'GO_Category') |> 
    dplyr::rename(cluster = 2, GO_Term = Description, GO_ID = ID) |> 
    tidyr::fill(cluster, .direction = 'down') |> 
    dplyr::filter(!is.na(GO_ID)) |> 
    dplyr::mutate(GO_Category = gsub(GO_Category,
                                     pattern = 'go_',
                                     replacement = '') %>% 
                    stringr::str_to_upper(.)) |> 
    tibble::remove_rownames() |> 
    dplyr::distinct()
  }
) %>%
  purrr::map_df(., .f = as.data.frame) |>  # join to the scores and hSAINT classes
  dplyr::full_join(cluster_scores) |> 
  dplyr::filter(!is.na(GO_Category)) |> 
  dplyr::inner_join(
    protein_cluster_df |> 
      dplyr::distinct(cluster, cluster_class)
  ) |> 
  dplyr::group_by(cluster) |> 
  dplyr::arrange(desc(mean_score), contam_perc, p.adjust)
```

```{r Load Cluster GO GSEA, eval = TRUE, echo = FALSE}
load("../../analysis/R/HDBSCAN/HDBSCANCluster_GO_GSEA.rda")
global_enrichment_results_df <- HDBSCANCluster_GO_GSEA
rm(HDBSCANCluster_GO_GSEA)

global_enrichment_results_df |> 
  dplyr::full_join(cluster_scores) |> 
  dplyr::filter(!is.na(GO_Category)) |> 
  dplyr::inner_join(
    protein_cluster_df |>
      dplyr::distinct(cluster, cluster_class)
  ) |>
  dplyr::group_by(cluster) |> 
  dplyr::arrange(desc(mean_score), contam_perc, p.adjust)
```

The "`High SAINT score`" clusters were among the clusters with significant 
enrichment of all the proteins detected in the experiments

However, many clusters (**30**) did not have significant enrichment relative to 
the entire set of proteins in the experiments

**TMEM is clustered into the noise cluster**, likely because of its low-variance
counts across treatments and weak correlation with other protein abundances; this 
is not uncommon for a transmembrane protein with no prior for potential interactors

Using `hierarchical SAINT score` as a explanatory tool, this is likely because 
the clusters without GO enrichment contained a large mix of proteins scored as 
"noise" (low posterior probability of being a true interactor) and signal 
(high posterior probability of being a true interactor)

In order to better make sense of the clustering-- describing the **"un-enriched" clusters**-- 
"sub-cluster" GO GSEA was conducted

## SubCluster GO GSEA

Sub-cluster enrichment analysis on clusters without any enrichment (to describe
themes of proteins within these clusters)

+ To do this, we adapt the method from hierarchical clustering vignette, where a 
dendrogram is created from the raw counts matrix

+ Then we create a local hierarchical tree and index it by filtering *for* the 
proteins that were not in an enriched cluster

```{r SubCluster index}
"%notin%" <- Negate("%in%")

# index the clusters without significant enrichment to describe sub-clusters
tier2_clusters <- c(seq(1:48) |> as.character())[
  which(c(seq(1:48) |> as.character()) %notin% c(global_enrichment_results_df |> 
          dplyr::distinct(cluster) |> 
          dplyr::pull()) == T)] 
```

```{r SubCluster function}
# function to gather sub-clusters for each cluster that did not return enrichment results
subclusters_from_hdbscan <- function(
  counts_mat,
  cluster_df,
  n_sub_per_cluster = 3L
) {
  # extract unique HDBSCAN clusters (exclude noise = 0)
  cluster_ids <- cluster_df |> 
    dplyr::filter(cluster != 'noise') |>  
    dplyr::distinct(cluster) |> dplyr::pull()

  # list to store subcluster assignments
  out_list <- vector("list", length(cluster_ids))
  names(out_list) <- as.character(cluster_ids)

  for (k in cluster_ids) {
    # proteins in this cluster
    prots_k <- cluster_df$Protein[cluster_df$cluster == k]

    # subset the counts matrix
    sub_mat <- counts_mat[prots_k, , drop = FALSE]

    # build a local hierarchical tree
    dist_obj <- stats::dist(sub_mat)
    hc_obj   <- stats::HDBSCAN(dist_obj, method = "ward.D2")

    # number of subclusters
    K_big <- n_sub_per_cluster

    # finer cut
    sub_labels <- stats::cutree(hc_obj, k = K_big)

    # assemble tibble
    df_k <- tibble::tibble(
      Protein   = prots_k,
      cluster   = k,
      sub_label = sub_labels
    )

    # define subcluster names
    df_k <- df_k %>%
      dplyr::group_by(cluster, sub_label) %>%
      dplyr::mutate(
        subcluster = paste0("sub", dplyr::cur_group_id())
      ) %>%
      dplyr::ungroup()

    out_list[[as.character(k)]] <- df_k
  }

  # combine all clusters
  dplyr::bind_rows(out_list)
}
```

```{r SubCluster application}
# get the sub-clusters
subclusters_df <- subclusters_from_hdbscan(
  counts_mat  = IPMS_mat,
  cluster_df = dplyr::filter(clusters_df, cluster %in% tier2_clusters),
  n_sub_per_cluster = 3L
)
```

```{r Show GSEA on SubClusters, eval= FALSE, echo = TRUE}
# perform GO GSEA on the subclusters

subcluster_enrichment_results <- subclusters_df %>%
  dplyr::inner_join(EntrezIDs |> dplyr::rename(Protein = 1)) |> 
  dplyr::group_by(cluster, subcluster) %>%
  dplyr::group_modify(~ {

    # subcluster genes
    genes <- .x$ENTREZID
    
    background_genes <- unique(subclusters_df$ENTREZID[subclusters_df$cluster == .y$cluster])

    tibble::tibble(
      BP = list(get_GO(genes, ontology = "BP", background_genes = background_genes)),
      CC = list(get_GO(genes, ontology = "CC", background_genes = background_genes)),
      MF = list(get_GO(genes, ontology = "MF", background_genes = background_genes))
    )
  }) %>%
  dplyr::ungroup()
rm(background_genes)
```

```{r Gather SubCluster GSEA data, eval = FALSE, echo = TRUE}
# wrangle the sub-cluster GO GSEA results
subcluster_enrichment_results_df <- subcluster_enrichment_results |> 
    tidyr::pivot_longer(
      cols      = c('BP', 'CC', 'MF'),
      names_to  = "GO_Category",
      values_to = "values"
    ) %>%
    dplyr::mutate(
      GO_tbl = purrr::map(values, ~ as.data.frame(.x))
    ) |>
    tidyr::unnest(GO_tbl) |>
    dplyr::rename(
      GO_ID   = ID,
      GO_Term = Description
    ) |>
    dplyr::select(
      cluster,
      subcluster,
      GO_Category,
      GO_Term,
      GO_ID,
      dplyr::everything(),
      -values
    )
```

```{r Load SubCluster GO_GSEA data, echo = F}
load("../../analysis/R/HDBSCAN/HDBSCANSubCluster_GO_GSEA.rda")
sublcuster_enrichment_results_df <- HDBSCANSubCluster_GO_GSEA
rm(HDBSCANCluster_GO_GSEA)
```

```{r Find UnEnriched SubClusters}
clusters_with_subcluster_enrichment <- tier2_clusters
clusters_without_subcluster_enrichment <- generics::setdiff(clusters_with_subcluster_enrichment,
                                                          subcluster_enrichment_results_df |> 
                                                            dplyr::distinct(cluster) |> 
                                                            dplyr::pull())
clusters_with_subcluster_enrichment <- subcluster_enrichment_results_df |> 
  dplyr::distinct(cluster) |> 
  dplyr::pull()
rm(tier2_clusters)
```

## Construct the Cluster GO (similarity) Network

Because `HDBSCAN` produced 47 clusters and a large number of enriched GO terms,
we applied semantic similarity reduction to collapse "redundant" GO terms prior 
to network visualizaation

Full GO results are retained in the accompanying tables

```{r Construct GO Network Object}
go_network <- global_enrichment_results_df |> 
  dplyr::full_join(subcluster_enrichment_results_df) |> 
  dplyr::full_join(cluster_scores) |> 
  dplyr::mutate(Tier = dplyr::case_when(is.na(subcluster) ~ 'Tier1',TRUE~'Tier2'),
                node_to = GO_Term,
                node_from = paste0(
                  
                  "cluster", cluster,
                  ifelse(is.na(subcluster),
                         '',
                         paste0('_Subcluster',
                                subcluster %>% 
                                  stringr::str_extract(., pattern = '\\d+')))),
                weight = -log10(p.adjust)) |> 
  dplyr::filter(weight > -log10(0.05))

g <- igraph::graph_from_data_frame(
  d = go_network[, c("node_from", "node_to",
                     "weight", "geneID", "p.adjust", "FoldEnrichment")],
  directed = TRUE,
  vertices =
    dplyr::bind_rows(
      go_network |> dplyr::transmute(name = node_from, score_mass, subcluster),
      go_network |> dplyr::transmute(name = node_to,
                                     score_mass = NA_real_,
                                     subcluster = NA_character_)
    ) |>
    dplyr::distinct(name, .keep_all = TRUE)
)
```

```{r, results='asis', echo=FALSE}
fmt <- knitr::pandoc_to()

if (identical(fmt, "html")) {
  cat('# Plots {.tabset .tabset-pills .tabset-fade}\n')
} else if (identical(fmt, "markdown")) {
  cat('<details><summary><strong>Plots</strong></summary>\n')
} else {
  cat('# Plots\n')
}
```

## Cluster Dot Plot

```{r RawClusterSubClusterDotPlotOfGOTerms, warning = F, message = F, fig.height=18}
global_enrichment_results_df |> 
  dplyr::full_join(subcluster_enrichment_results_df) |> 
  dplyr::full_join(cluster_scores) |> 
  dplyr::filter(!is.na(GO_Term)) |> 
  dplyr::filter(!is.na(GO_Category)) |> 
  dplyr::mutate(
    Tier = dplyr::case_when(
      is.na(subcluster) ~ paste0('Clust. ', cluster),
      TRUE ~ paste0('Clust. ', cluster,
                    ':\nsubclust. ',
                    gsub(subcluster, pattern = 'sub', replacement = ''))
      ),
    is_enriched = dplyr::case_when(is.na(subcluster) ~ 'Enriched',
                                   TRUE ~ 'Not Enriched')
    ) |> 
  dplyr::group_by(cluster) |>  
  dplyr::slice_head(n = 3) |>
  dplyr::arrange(p.adjust) |>
  ggplot2::ggplot(
    ggplot2::aes(
      x = -log10(p.adjust),
      y = forcats::fct_rev(forcats::fct_infreq(GO_Term, mean_score)),
      color = GO_Category,
      size = FoldEnrichment,
      alpha = mean_score,
      shape = is_enriched,
    )
  ) +
  ggplot2::geom_vline(xintercept = -log10(0.05),
                      linetype = 'dashed',
                      color = 'firebrick',
                      size = 1,
                      alpha = 0.5) +
  ggplot2::geom_point() +
  ggplot2::facet_grid(Tier ~ ., scales = 'free_y') +
  ggplot2::labs(title = 'GO Enrichment Across\nClusters and SubClusters\nin TMEM184B IPMS\nHierarchical Clustering',
                x = '-log10(Adj.P-val)',
                y = 'GO Term',
                alpha = 'Cluster\nMean\nHierarchical\nSAINT\nScore',
                size = 'FoldEnrichment',
                color = 'GO Category',
                shape = 'IsEnrichedcluster') +
  ggplot2::theme_bw() +
  ggplot2::theme(title = ggplot2::element_text(face = 'bold', size = 12),
                 axis.text = ggplot2::element_text(face = 'bold', size = 10),
                 strip.text = ggplot2::element_text(face = 'bold', size = 7), 
                 strip.background = ggplot2::element_rect(color = 'black',
                                                          fill = 'white')) +
  ggplot2::guides(#shape = ggplot2::guide_legend(override.aes = list(size = 4, shape = c(19, 17))),
                  color = ggplot2::guide_legend(override.aes = list(size = 4,
                                                                    shape = 19,
                                                                    color = c('navy',
                                                                              'darkgoldenrod3',
                                                                              'firebrick'))),
                  alpha = ggplot2::guide_legend(override.aes = list(size = 5,
                                                                    shape = 19)),
                  size = ggplot2::guide_legend(override.aes = list(shape = 19))
                  # size = 'none'
                  ) +
  ggplot2::scale_color_manual(values = c('navy',
                                        'darkgoldenrod3',
                                        'firebrick'))
```

## Cluster GO (similarity) Network

```{r ClusterGOTermSimilarityNetwork, warning = F, message = F}
coords <- igraph::layout_with_fr(g)
coords <- coords*200

visNetwork::visNetwork(
  nodes = data.frame(
    id    = igraph::V(g)$name,
    label = igraph::V(g)$name,
    size = scales::rescale(
      (\(x){ x[is.na(x)] <- 0; x })(igraph::V(g)$score_mass),
      to = c(20, 60)
    ),
    x = coords[,1],
    y = coords[,2],
    group = base::interaction(
                        {
                          x <- igraph::V(g)$cluster
                          if (length(x) == 0) rep("", igraph::vcount(g)) else x
                        },
                        {
                          x <- igraph::V(g)$subcluster
                          if (length(x) == 0) {
                            rep("", igraph::vcount(g))
                          } else {
                            ifelse(nzchar(x), x, igraph::V(g)$cluster)
                          }
                        },
                        drop = TRUE,
                        lex.order = TRUE
                      ),
    title = sapply(
      igraph::V(g),
      function(v) {
        incoming_edges <- igraph::incident(g, v, mode = "all")
        incoming_edges <- incoming_edges[
          igraph::ends(g, incoming_edges)[,2] == igraph::V(g)[v]$name
        ]
        raw_gene_strings <- igraph::E(g)[incoming_edges]$geneID
        raw_gene_strings <- unlist(raw_gene_strings, use.names = FALSE)

        genes <- raw_gene_strings |>
          as.character() |>
          (\(x) x[!is.na(x)])() |>
          strsplit("\\\\") |>
          unlist() |>
          (\(x) x[nzchar(x)])() |>
          unique()

        term <- igraph::V(g)[v]$name
        saint_value <- igraph::V(g)[v]$score_mass
        if (is.na(saint_value)) saint_value <- 0
        is_module <- nzchar(igraph::V(g)[v]$subcluster)

        if (is_module) {
          paste0(
            "<div style='background-color:#ffffff;color:#1f4e79;",
            "padding:6px 10px;border-radius:6px;",
            "border:1px solid #1f4e79;font-size:12px;'>",
            "<b>", term, "</b><br>",
            "<b>SAINT mass:</b> ", signif(saint_value, 3), "<br>",
            if (length(genes) > 0)
              paste0("<u>Proteins:</u><br>", paste(genes, collapse = "<br>"))
            else "",
            "</div>"
          )
        } else {
          paste0(
            "<div style='background-color:#ffffff;color:#1f4e79;",
            "padding:6px 10px;border-radius:6px;",
            "border:1px solid #1f4e79;font-size:12px;'>",
            "<b>", term, "</b><br>",
            if (length(genes) > 0)
              paste0("<u>Proteins:</u><br>", paste(genes, collapse = "<br>"))
            else "",
            "</div>"
          )
        }
      }
    )
  ),
  edges = data.frame(
    from  = igraph::as_data_frame(g, "edges")$from,
    to    = igraph::as_data_frame(g, "edges")$to,
    value = igraph::E(g)$weight,
    title = vapply(
      seq_along(igraph::E(g)),
      function(i) {
        paste0(
          "<div style='background-color:#ffffff;color:#1f4e79;",
          "padding:6px 10px;border-radius:6px;",
          "border:1px solid #1f4e79;font-size:12px;'>",
          "<b>p.adjust:</b> ", signif(igraph::E(g)$p.adjust[i], 3), "<br>",
          "<b>weight:</b> ", signif(igraph::E(g)$weight[i], 3),
          "</div>"
        )
      },
      FUN.VALUE = character(1)
    )
  )
) |>
  # visNetwork::visNodes(color = NULL) |>
  visNetwork::visEdges(arrows = "to") |>
  visNetwork::visNodes(
    scaling = list(min = 5, max = 40),
    shadow  = TRUE
  ) |>
  visNetwork::visLayout(randomSeed = 1) |>
  visNetwork::visPhysics(enabled = FALSE) |>
  visNetwork::visInteraction(
    dragView = TRUE,
    zoomView = TRUE
  ) |>
  visNetwork::visOptions(nodesIdSelection = TRUE)
rm(coords)
```

Clusters with `enriched` GO Terms are `blue`

Here, only the enriched sub-clusters were included:
the clusters with that did not reach significant enrichment broadly, but did 
reach significant enrichment when that cluster was broken up. These `Enriched` 
**sub-clusters** are **non-blue**

```{r, results='asis', echo=FALSE}
fmt <- knitr::pandoc_to()

if (identical(fmt, "html")) {
  cat('# Results Tables {.tabset .tabset-pills .tabset-fade}\n')
} else if (identical(fmt, "markdown")) {
  cat('<details><summary><strong>Results Tables</strong></summary>\n')
} else {
  cat('# Results Tables\n')
}
```

```{r restore-shiny, include=FALSE}
options(htmlwidgets.shinyMode = TRUE)
```

## Clustering Results Table by Protein (Clusters 9, 12)

```{r Print Protein Level Results Table}
DT::datatable(
  protein_cluster_df |> 
    dplyr::full_join(global_enrichment_results_df) |> 
    dplyr::full_join(subcluster_enrichment_results_df) |> 
    dplyr::rename(hSAINT_score = score) |> 
    dplyr::rename(cluster_size = n) %>%
    dplyr::rename_with(.cols = c(dplyr::contains("mean_"),
                                 dplyr::contains("median_"),
                                 dplyr::contains("_perc"),
                                 score_mass), ~paste0("cluster_", .x)) |> 
    dplyr::filter(!is.na(GO_Term)) |> 
    dplyr::filter(!is.na(GO_Category)) |> 
    dplyr::filter( grepl(classification,
                            pattern = 'High SAINT, Low') &
                     hSAINT_score > 0.5
                   )
)
```

This (complete) table exists as [HDBSCAN_protein_level_GO_GSEA.rda](../../analysis/R/HDBSCAN/HDBSCAN_protein_level_GO_GSEA.rda) in [Endo/analysis/R/HDBSCAN/](../../analysis/R/HDBSCAN)

## Clustering Results Table by Cluster

```{r Print Cluster Level Results Table}
DT::datatable(
  global_enrichment_results_df |> 
  dplyr::full_join(subcluster_enrichment_results_df) |> 
  dplyr::full_join(cluster_scores) |> 
  dplyr::rename(cluster_size = n) %>%
  dplyr::rename_with(.cols = c(dplyr::contains("mean_"),
                                 dplyr::contains("median_"),
                                 dplyr::contains("_perc"),
                                 score_mass), ~paste0("cluster_", .x)) |> 
  dplyr::filter(!is.na(GO_Term)) |> 
  dplyr::filter(!is.na(GO_Category))
)
```

This table exists as [HDBSCAN_cluster_level_GO_GSEA.rda](../../analysis/R/HDBSCAN/HDBSCAN_cluster_level_GO_GSEA.rda) in [Endo/analysis/R/HDBSCAN/](../../analysis/R/HDBSCAN)

```{r, results='asis', echo=FALSE}
fmt <- knitr::pandoc_to()

if (identical(fmt, "html")) {
  cat('# Summary {.tabset .tabset-pills .tabset-fade}\n')
} else if (identical(fmt, "markdown")) {
  cat('<details><summary><strong>Summary</strong></summary>\n')
} else {
  cat('# Summary\n')
}
```
