---
title: "Hierarchical Clustering vignette"
author: "Erik Larsen"
date: "`r Sys.Date()`"
always_allow_html: true
output:
  html_document:
    toc: true
    fig_width: 10
    fig_height: 10
  pdf_document:
    fig_width: 10
    fig_height: 10
    latex_engine: xelatex
  md_document:
    variant: gfm
vignette: >
  %\VignetteIndexEntry{Hierarchical Clustering vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Overview

This markdown was developed to formally document an ad-hoc cluster analysis for 
TMEM184B immunoprecipitation tandem mass-spec experiments for 
**Dr. Martha Bhattacharya's lab** at the **University of Arizona** for the
endolysosomal paper under revision (2025-02-28) for the Journal of Cell Science

This vignette documents the `Hierarchical Clustering` of this IPMS data

```{r reset-shiny-mode, include=FALSE}
options(htmlwidgets.shinyMode = FALSE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.height = 12,
  dev = c("png", "pdf"),
  fig.path = "../../analysis/R/hClust/"
)
```

## Attach Packages

```{r Attach Packages, message=F,warning=F}
# data/wrangling
library(tidyverse)
library(Endo)
library(DT)

# clustering
library(pheatmap)
library(dynamicTreeCut)

# GSEA/viz
library(clusterProfiler)
library(org.Hs.eg.db)
library(ReactomePA)
library(AnnotationDbi)

# progress bar
library(progressr)

# parallelization
library(future)
library(future.apply)

# network viz
library(visNetwork)

options(stringsAsFactors = FALSE)
```

## Load the Pre-Computed Data

```{r Data Load}
# Load precomputed objects from preprocessing
load("../../analysis/R/precompute_objects.rda")
```

```{r Second Data Load, echo = FALSE}
rm(IPMS_mat_log)
hSAINT <- hSAINT |> 
  dplyr::mutate(protein_stat_class = dplyr::case_when(score > 0.5 & 
                                                lit_flag + CRAPome_flag < 1 ~ 'High hSAINT, Low Contamination',
                                              score < 0.5 &
                                                lit_flag + CRAPome_flag < 1 ~ 'Low hSAINT, Low Contamination',
                                              score > 0.5 &
                                                lit_flag + CRAPome_flag > 0 ~ 'High hSAINT, High Contamination',
                                              score < 0.5 &
                                                lit_flag + CRAPome_flag > 0 ~ 'Low hSAINT, High Contamination'),
                contam_flag = lit_flag + CRAPome_flag) |> 
  dplyr::rename(protein_hSAINT_score = score) |> 
  dplyr::rename_with(.cols = c(dplyr::contains("flag")), ~paste0('protein_', .x)) |> 
  dplyr::relocate(protein_contam_flag, .after = protein_hSAINT_score)
```


```{r, results='asis', echo=FALSE}
fmt <- knitr::pandoc_to()

if (identical(fmt, "html")) {
  cat('# Hierarchical Clustering {.tabset .tabset-pills .tabset-fade}\n')
} else if (identical(fmt, "markdown")) {
  cat('<details><summary><strong>Hierarchical Clustering</strong></summary>\n')
} else {
  cat('# Hierarchical Clustering\n')
}
```

## Raw Counts

```{r Hierarchical Clustering raw}
d <- stats::as.dist(1 - stats::cor(t(IPMS_mat), use = 'pairwise.complete.obs'))
hc <- stats::hclust(d, method = "average")
```

```{r Grid Search Cluster Size}
sizes <- c(20, 30, 50, 80, 100)
results <- list()

for (m in sizes) {
  cl <- dynamicTreeCut::cutreeDynamic(
    dendro = hc,
    distM = as.matrix(d),
    deepSplit = 1,
    pamRespectsDendro = TRUE,
    minClusterSize = m
  )
  results[[as.character(m)]] <- cl
  cat("minClusterSize =", m, "→", length(table(cl)), "clusters\n")
}
rm(list = c('cl', 'm', 'sizes', 'results'))
```

  + `optimal minimum cluster size` is **50** (elbow; diminishing returns after)
  
  + `# clusters` = **28**

```{r Clusters on Raw Counts}
clusters_opt <- dynamicTreeCut::cutreeDynamic(hc,
                                              distM = as.matrix(d),
                                              deepSplit = 1,
                                              pamRespectsDendro = T,
                                              minClusterSize = 50)
names(clusters_opt) <- rownames(IPMS_mat)
```

```{r Get Clusters}
cluster_ids <- sort(unique(clusters_opt))
cluster_colors <- stats::setNames(
  grDevices::rainbow(length(cluster_ids), s = 0.6, v = 0.9),
  cluster_ids
)

annotation_colors <- list(cluster = cluster_colors)
```

Sparse data is too difficult to discern from a global view

  -> Z transform the raw counts **after clustering** to plot
  
  + the cluster object preserves the ordering of the proteins (dendrogram)
  
    + Z-score therefore only changes the values of the columns and rows to 
    better visualize the sparse data
  
```{r Scale the Matrix Used for Clustering by Zscore}
IPMS_z <- t(scale(t(IPMS_mat)))
colnames(IPMS_z) <- colnames(IPMS_mat)

condition <- sub('_.*', '', colnames(IPMS_mat))

col_order <- unlist(
  lapply(
    split(seq_along(condition), condition), function(idx) {
      submat <- IPMS_z[, idx, drop = FALSE]
      hc_sub <- stats::hclust(stats::dist(t(submat)), method = 'ward.D2')
      idx[hc_sub$order]
    }))

IPMS_z_ordered <- IPMS_z[, col_order]

rm(list = c('condition', 'col_order'))
```

```{r, results='asis', echo=FALSE}
fmt <- knitr::pandoc_to()

if (identical(fmt, "html")) {
  cat('# Cluster Descriptions {.tabset .tabset-pills .tabset-fade}\n')
} else if (identical(fmt, "markdown")) {
  cat('<details><summary><strong>Cluster Descriptions</strong></summary>\n')
} else {
  cat('# Cluster Descriptions\n')
}
```
In order to assess this method and also to compare/validate the hierarchical SAINT
model, describing clusters based on the proteins in them is useful

Similarly localized proteins could be expected to have similar abundances to 
each other as well as the target (**TMEM184B**)

Similar functioning proteins could also be expected to have similar abundances 
to each other as well as the target

Describing these clusters attained by `hierarchical clustering` via 
gene ontology gene set enrichment analysis is, therefore, useful and standard 
practice

Furthermore, we may be able to assemble a `protein-protein interaction network`
from the results

I outline this below-- the methods and results for acquiring these clusters, 
their constituent proteins, and how we make sense of them-- visualized as
`dotplots`, `heatmaps`, and a `network`. I include searchable tables printed last

## Extract Clusters

Extract gene names from the hierarchically-clustered clusters and convert to a 
dataframe

```{r Extract Gene Names from Clusters}
clusters_list_idx <- lapply(sort(unique(clusters_opt)), function(m) {
  which(clusters_opt == m)
})
names(clusters_list_idx) <- sort(unique(clusters_opt))

clusters_list <- lapply(sort(unique(clusters_opt)), function(m) {
  rownames(IPMS_mat)[clusters_opt == m]
})
names(clusters_list) <- sort(unique(clusters_opt)) |> as.character()

clusters_df <- clusters_list %>%
  purrr::map_df(., .f = as.data.frame, .id = 'cluster') |> 
  dplyr::rename(Protein = 2) |> 
  dplyr::relocate(Protein, .before = 1)
```

```{r, echo = FALSE}
rm(list = ls(pattern = '_mat'))
```

## Join Clusters to hSAINT

Include score summaries (mean and median posterior probability scores, ranks, and literature or CRAPome contamination rate)

```{r Summarize SAINT Scores and Contamination Rates}
cluster_scores <- clusters_df |>  
  dplyr::inner_join(
    hSAINT
  ) |> 
  dplyr::arrange(desc(protein_hSAINT_score)) |> 
  dplyr::group_by(cluster) |> 
  dplyr::summarize(cluster_mean_hSAINT_score = mean(protein_hSAINT_score),
                   cluster_median_hSAINT_score = stats::median(protein_hSAINT_score),
                   cluster_hSAINT_score_mass = sum(protein_hSAINT_score),
                   cluster_size = dplyr::n(),
                   cluster_lit_flag_perc = (sum(protein_lit_flag)/cluster_size)*100,
                   cluster_CRAPome_flag_perc = (sum(protein_CRAPome_flag)/cluster_size)*100,
                   cluster_contam_perc = cluster_lit_flag_perc + cluster_CRAPome_flag_perc) |> 
  dplyr::arrange(desc(cluster_mean_hSAINT_score), cluster_contam_perc) |> 
  dplyr::mutate(cluster_mean_rank = dplyr::row_number()) |> 
  dplyr::arrange(desc(cluster_median_hSAINT_score), cluster_contam_perc) |> 
  dplyr::mutate(cluster_median_rank = dplyr::row_number()) |> 
  dplyr::arrange(desc(cluster_hSAINT_score_mass), cluster_contam_perc) |> 
  dplyr::mutate(cluster_total_rank = dplyr::row_number()) |> 
  dplyr::arrange(cluster_mean_rank) |> 
  dplyr::relocate(c(cluster_lit_flag_perc, cluster_CRAPome_flag_perc), .after = cluster_total_rank)
```

Join the cluster scores to the cluster (per protein) dataframe and the `hSAINT` results for classification

```{r Join the Summary dataframe to the Protein Level dataframes}
protein_cluster_df <- clusters_df |> 
  dplyr::inner_join(
    hSAINT
  ) |> 
  dplyr::full_join(
    cluster_scores |> 
      dplyr::ungroup()
  ) |> 
  dplyr::mutate(dplyr::across(c(dplyr::where(is.double)), ~ round(.x, 3))) |> 
  dplyr::mutate(
    cluster_class = dplyr::case_when(
      cluster_median_hSAINT_score > 0.5 &
        cluster_contam_perc > 
        cluster_scores |> 
        dplyr::pull(cluster_contam_perc) |> 
        stats::median() ~ 'High hSAINT, High Contamination',
      cluster_median_hSAINT_score > 0.5 &
        cluster_contam_perc <
        cluster_scores |> 
        dplyr::pull(cluster_contam_perc) |> 
        stats::median() ~ 'High hSAINT, Low Contamination',
      cluster_median_hSAINT_score < 0.5 &
        cluster_contam_perc > 
        cluster_scores |> 
        dplyr::pull(cluster_contam_perc) |> 
        stats::median() ~ 'Low hSAINT, High Contamination',
      cluster_median_hSAINT_score < 0.5 &
        cluster_contam_perc <
        cluster_scores |> 
        dplyr::pull(cluster_contam_perc) |> 
        stats::median() ~ 'Low hSAINT, Low Contamination'
    )) |> 
  dplyr::relocate(cluster_class, .after = cluster) |> 
  dplyr::arrange(desc(protein_hSAINT_score), Protein) |> 
  dplyr::group_by(cluster) |> 
  dplyr::mutate(cluster_clean_n = sum(protein_contam_flag == 0), .after = cluster_size) |> 
  dplyr::mutate(cluster_clean_fraction = round((cluster_clean_n/cluster_size)*100, 2), .after = cluster_clean_n) |> 
  dplyr::ungroup()

head(protein_cluster_df)
```

## Cluster GO GSEA

```{r Downstream Analysis}
EntrezIDs <- AnnotationDbi::select(
  org.Hs.eg.db,
  keys = clusters_df |> dplyr::distinct(Protein) |> dplyr::pull(),
  columns = 'ENTREZID',
  keytype = 'SYMBOL'
) |> 
  dplyr::filter(!is.na(2))

EntrezMap <- stats::setNames(EntrezIDs$ENTREZID, EntrezIDs$SYMBOL)
```

Create enrichment functions

```{r Create Enrichment Functions}
background_genes <- EntrezIDs |> # use the EntrezIDs mapped from the modules_df
  dplyr::pull(ENTREZID)

get_GO <- function(genes, ontology, background_genes) {
  clusterProfiler::enrichGO(
    gene = genes,
    universe = background_genes,
    OrgDb = org.Hs.eg.db,
    keyType = 'ENTREZID',
    ont = ontology,
    pAdjustMethod = 'BH',
    qvalueCutoff = 0.05,
    readable = T
  )
}

get_kegg <- function(genes, EntrezMap) {
  entrez <- EntrezMap[genes]
  entrez <- entrez[!is.na(entrez)]
  if (length(entrez)==0) {
    return(NULL)
  }
  clusterProfiler::enrichKEGG(
    gene = entrez,
    organism = 'hsa',
    pAdjustMethod = 'BH',
    qvalueCutoff = 0.05
  )
}

get_reactome <- function(genes, EntrezMap) {
  entrez <- EntrezMap[genes]
  entrez <- entrez[!is.na(entrez)]
  if (length(entrez)==0) {
    return(NULL)
  }
  ReactomePA::enrichPathway(
    gene = entrez,
    organism = 'human',
    pAdjustMethod = 'BH',
    qvalueCutoff = 0.05,
    readable = T
  )
}
```

Run the functions on the dataset in parallel (multiple CPUs)

```{r Parallelize Enrichment, eval = FALSE}
future::plan(future::multisession, workers = 6)

progressr::with_progress({

  cluster_ids <- sort(unique(clusters_df$cluster))
  
  p <- progressr::progressor(steps = length(cluster_ids))

    enrichment_results <- future.apply::future_lapply(
      cluster_ids,
      function(mod) {
        
        # tick progress IN THE MAIN SESSION
      p(sprintf("cluster %s", mod))
    
        suppressPackageStartupMessages({
          library(clusterProfiler)
          library(DOSE)
          library(org.Hs.eg.db)
          library(dplyr) 
        })
    
        get_GO <- get_GO   # export the function so it's recognized for parallelization
        protein_cluster_df <- protein_cluster_df
        clusters_df <- clusters_df
    
        # proteins without any contamination except for one cluster
        genes <- clusters_df |> 
          dplyr::filter(cluster == mod) |> 
          dplyr::inner_join(EntrezIDs |> 
                              dplyr::rename(Protein = 1)) |> # get EntrezIDs
          dplyr::inner_join(protein_cluster_df |> 
                              dplyr::filter(protein_contam_flag == 0) |>
                              dplyr::distinct(Protein)) |> 
          dplyr::distinct(ENTREZID) |> 
          dplyr::pull()
        
        # all proteins within a given cluster
        background_genes <- clusters_df |> 
          dplyr::filter(cluster == mod) |> 
          dplyr::inner_join(EntrezIDs |> 
                              dplyr::rename(Protein = 1)) |> 
          dplyr::distinct(ENTREZID) |> 
          dplyr::pull()
    
        list(
          cluster = mod,
          go_bp  = get_GO(genes, ontology = "BP", background_genes = background_genes),
          go_cc  = get_GO(genes, ontology = "CC", background_genes = background_genes),
          go_mf  = get_GO(genes, ontology = "MF", background_genes = background_genes)
        )
        }
    ) 
})

#names(enrichment_results) <- cluster_ids
rm(p)
```

Gather enrichment results

```{r Consolidate Enrichment Results, eval = FALSE}
global_enrichment_results_df <- purrr::map_dfr(
  enrichment_results,
  function(x) {

    out <- list()

    for (ont in c("BP", "CC", "MF")) {

      slot_name <- paste0("go_", tolower(ont))

      res <- x[[slot_name]]

      if (!is.null(res)) {
        df <- res@result
        if (is.data.frame(df)) {
          if (nrow(df) > 0) {

            df$cluster <- x$cluster
            df$GO_Category <- ont

            out[[length(out) + 1]] <- df
          }
        }
      }
    }

    if (length(out) == 0) {
      return(NULL)
    } else {
      return(dplyr::bind_rows(out))
    }
  }
) |>
  dplyr::rename(GO_ID = ID, GO_Term = Description) |>
  dplyr::relocate(c(cluster, GO_Category), .before = 1) |>
  tibble::rowid_to_column("id") |>
  dplyr::filter(!is.na(GO_Category)) |>
  dplyr::select(-id) |>
  dplyr::full_join(
    cluster_scores
  ) |>
  dplyr::inner_join(
    protein_cluster_df |> dplyr::distinct(cluster, cluster_class)
  ) |>
  dplyr::group_by(cluster) |>
  dplyr::arrange(cluster_contam_perc, desc(cluster_mean_hSAINT_score), p.adjust) |>
  dplyr::filter(-log10(p.adjust) >= -log10(0.05)) |> 
  dplyr::arrange(cluster_contam_perc, desc(p.adjust)) |> 
  dplyr::relocate(cluster_class, .after = 2)
```

```{r Load Cluster GO GSEA, eval = TRUE, echo = FALSE}
load("../../analysis/R/hClust/HierarchicalCluster_GO_GSEA.rda")
global_enrichment_results_df <- HierarchicalCluster_GO_GSEA
rm(HierarchicalCluster_GO_GSEA)
```

The "`High SAINT score`" clusters were among the clusters with significant 
enrichment of all the proteins detected in the experiments

However, many clusters did not have significant enrichment relative to the entire
set of proteins in the experiments, *including TMEM's*. Using `hierarchical SAINT score` 
as a explanatory tool, this is likely because the clusters without GO enrichment 
contained a large mix of proteins scored as "noise" (low posterior probability 
of being a true interactor) and signal (high posterior probability of being a 
true interactor)

In order to better make sense of the clustering-- describing the **"un-enriched" clusters**-- 
"sub-cluster" GO GSEA was conducted

## SubCluster GO GSEA

Sub-cluster enrichment analysis on clusters without any enrichment (to describe
themes of proteins within these clusters)

```{r SubCluster index}
"%notin%" <- Negate("%in%")

# index the clusters without significant enrichment to describe sub-clusters
tier2_clusters <- c(seq(1:28) |> as.character())[
  which(c(seq(1:28) |> as.character()) %notin% c(global_enrichment_results_df |> 
          dplyr::distinct(cluster) |> 
          dplyr::pull()) == T)] 
```

```{r SubCluster function}
# function to gather sub-clusters for each cluster that did not return enrichment results
subclusters_from_hclust <- function(
  hc_object,
  clusters_df,
  n_sub_per_cluster = 3L
) {
  # number of existing clusters
  n_clusters <- length(unique(clusters_df$cluster))

  # total number of clusters in the finer cut
  K_big <- n_clusters * n_sub_per_cluster

  # finer cut of the same tree
  sub_labels_all <- stats::cutree(hc_object, k = K_big)

  # assemble into a tibble
  sub_df <- tibble::tibble(
    Protein   = names(sub_labels_all),
    sub_label = sub_labels_all
  )

  # attach existing cluster labels
  sub_df <- dplyr::inner_join(
    sub_df,
    clusters_df,
    by = "Protein"
  )

  # define subclusters as (cluster, sub_label) pairs
  subclusters_df <- sub_df %>%
    dplyr::group_by(cluster, sub_label) %>%
    dplyr::mutate(
      subcluster = paste0("sub", dplyr::cur_group_id())
    ) %>%
    dplyr::ungroup() %>%
    dplyr::select(Protein, cluster, subcluster)

  return(subclusters_df)
}
```

```{r SubCluster application}
# get the sub-clusters
subclusters_df <- subclusters_from_hclust(
  hc_object  = hc,
  clusters_df = dplyr::filter(clusters_df, cluster %in% tier2_clusters),
  n_sub_per_cluster = 3L
)
```

```{r Show GSEA on SubClusters, eval= FALSE, echo = TRUE}
future::plan(future::multisession, workers = 6)

progressr::with_progress({

  sub_ids <- subclusters_df |>
    dplyr::distinct(cluster, subcluster) |>
    dplyr::arrange(cluster, subcluster)

  p <- progressr::progressor(steps = nrow(sub_ids))

  subcluster_enrichment <- future.apply::future_lapply(
    seq_len(nrow(sub_ids)),
    function(i) {

      cl <- sub_ids$cluster[i]
      sc <- sub_ids$subcluster[i]

      p(sprintf("cluster %s subcluster %s", cl, sc))

      suppressPackageStartupMessages({
        library(clusterProfiler)
        library(DOSE)
        library(org.Hs.eg.db)
        library(dplyr)
      })

      # attach relevant objects
      sub_ids <- sub_ids
      subclusters_df <- subclusters_df
      EntrezIDs <- EntrezIDs
      protein_cluster_df <- protein_cluster_df
      get_GO <- get_GO
      
      # retain only proteins that are not contaminants for enrichment
      genes <- subclusters_df |>
        dplyr::filter(cluster == cl, subcluster == sc) |>
        dplyr::inner_join(
          EntrezIDs |> 
            dplyr::filter(!is.na(ENTREZID)) |> 
            dplyr::rename(Protein = 1),
          by = "Protein"
        ) |>
        dplyr::inner_join(
          protein_cluster_df |>
            dplyr::filter(protein_contam_flag == 0) |>
            dplyr::distinct(Protein),
          by = "Protein"
        ) |>
        dplyr::distinct(ENTREZID) |>
        dplyr::pull()

      # use all proteins as background
      background_genes <- subclusters_df |>
        dplyr::filter(cluster == cl) |>
        dplyr::inner_join(
          EntrezIDs |> 
            dplyr::filter(!is.na(ENTREZID)) |> 
            dplyr::rename(Protein = 1),
          by = "Protein"
        ) |>
        dplyr::distinct(ENTREZID) |>
        dplyr::pull()

      out <- list()

      for (ont in c("BP", "CC", "MF")) {

        res <- get_GO(genes, ontology = ont, background_genes = background_genes)

        if (!is.null(res)) {
          if (nrow(res@result) > 0) {
            df <- as.data.frame(res)
            if (nrow(df) > 0) {
              df$cluster <- cl
              df$subcluster <- sc
              df$GO_Category <- ont
              out[[length(out) + 1]] <- df
            }
          }
        }
      }

      if (length(out) == 0) {
        return(NULL)
      } else {
        return(dplyr::bind_rows(out))
      }
    }
  )

  dplyr::bind_rows(subcluster_enrichment)
})
```

```{r Gather SubCluster GSEA data, eval = FALSE, echo = TRUE}
# wrangle the sub-cluster GO GSEA results

subcluster_enrichment_results_df <- subcluster_enrichment %>% 
  purrr::map_df(., .f = as.data.frame) |> 
  dplyr::rename(GO_ID = ID, GO_Term = Description) |> 
  dplyr::relocate(c(cluster, subcluster, GO_Category), .before = 1) |> 
  tibble::rowid_to_column('id') |> 
  dplyr::filter(!is.na(GO_Category)) |> 
  dplyr::select(-id) |> 
  dplyr::full_join(
    cluster_scores
  ) |>
  dplyr::inner_join(
    protein_cluster_df |> 
      dplyr::distinct(cluster, cluster_class)
  ) |> 
  dplyr::group_by(cluster) |> 
  dplyr::arrange(cluster_contam_perc, desc(cluster_median_hSAINT_score), p.adjust) |> 
  dplyr::filter(!is.na(GO_Category)) |> 
  dplyr::relocate(cluster_class, .after = 3)

subcluster_enrichment_results_df |> 
  head()
```

```{r Load SubCluster GO GSEA, eval = TRUE, echo = FALSE}
load("../../analysis/R/hClust/HierarchicalSubClustCluster_GO_GSEA.rda")
subcluster_enrichment_results_df <- HierarchicalSubClustCluster_GO_GSEA
rm(HierarchicalSubClustCluster_GO_GSEA)
```

```{r Find UnEnriched SubClusters}
clusters_with_subcluster_enrichment <- tier2_clusters
clusters_without_subcluster_enrichment <- generics::setdiff(clusters_with_subcluster_enrichment,
                                                          subcluster_enrichment_results_df |> 
                                                            dplyr::distinct(cluster) |> 
                                                            dplyr::pull())
clusters_with_subcluster_enrichment <- subcluster_enrichment_results_df |> 
  dplyr::distinct(cluster) |> 
  dplyr::pull()
rm(tier2_clusters)
```

## Construct the Cluster GO (similarity) Network

```{r Construct GO Network Object}
go_network <- global_enrichment_results_df |> 
  dplyr::full_join(subcluster_enrichment_results_df |> 
                     dplyr::rename_with(.cols = c(dplyr::contains("rank")), ~paste0("cluster_", .x))) |> 
  dplyr::mutate(Tier = dplyr::case_when(is.na(subcluster) ~ 'Tier1',TRUE~'Tier2'),
                node_to = GO_Term,
                node_from = paste0(
                  
                  "cluster", cluster,
                  ifelse(is.na(subcluster),
                         '',
                         paste0('_Subcluster',
                                subcluster %>% 
                                  stringr::str_extract(., pattern = '\\d+')))),
                weight = -log10(p.adjust)) |> 
  dplyr::filter(weight > -log10(0.05)) |> 
  dplyr::ungroup()

g <- igraph::graph_from_data_frame(
  d = go_network |> 
    dplyr::ungroup() |> 
    dplyr::transmute(
      from = node_from,
      to = node_to,
      weight,
      geneID,
      p.adjust,
      FoldEnrichment
    ),
  directed = TRUE,
  vertices = go_network |>
    dplyr::distinct(
      name = node_from,
      cluster,
      subcluster,
      cluster_hSAINT_score_mass
      ) |>
    dplyr::right_join(
      tibble::tibble(
        name = unique(c(go_network$node_from, go_network$node_to))
        ),
      by = "name"
      )
  )
```

```{r, results='asis', echo=FALSE}
fmt <- knitr::pandoc_to()

if (identical(fmt, "html")) {
  cat('# Plots {.tabset .tabset-pills .tabset-fade}\n')
} else if (identical(fmt, "markdown")) {
  cat('<details><summary><strong>Plots</strong></summary>\n')
} else {
  cat('# Plots\n')
}
```

## Cluster Dot Plot

```{r RawClusterSubClusterDotPlotOfGOTerms, warning = F, message = F}
global_enrichment_results_df |> 
  dplyr::full_join(subcluster_enrichment_results_df) |> 
  dplyr::full_join(cluster_scores) |> 
  dplyr::filter(!is.na(GO_Term)) |> 
  dplyr::filter(!is.na(GO_Category)) |> 
  dplyr::mutate(
    Tier = dplyr::case_when(
      is.na(subcluster) ~ paste0('Mod. ', cluster),
      TRUE ~ paste0('Mod. ', cluster,
                    ':\nsubmod. ',
                    gsub(subcluster, pattern = 'sub', replacement = ''))
      ),
    is_enriched = dplyr::case_when(is.na(subcluster) ~ 'Enriched',
                                   TRUE ~ 'Not Enriched')
    ) |> 
  dplyr::group_by(cluster) |>  
  dplyr::arrange(p.adjust) |>
  ggplot2::ggplot(
    ggplot2::aes(
      x = -log10(p.adjust),
      y = forcats::fct_rev(forcats::fct_infreq(GO_Term, cluster_median_hSAINT_score)),
      color = GO_Category,
      size = FoldEnrichment,
      alpha = cluster_median_hSAINT_score,
      shape = is_enriched,
    )
  ) +
  ggplot2::geom_vline(xintercept = -log10(0.05),
                      linetype = 'dashed',
                      color = 'firebrick',
                      size = 1,
                      alpha = 0.5) +
  ggplot2::geom_point() +
  ggplot2::facet_grid(Tier ~ ., scales = 'free_y') +
  ggplot2::labs(title = 'GO Enrichment Across\nClusters and SubClusters\nin TMEM184B IPMS\nHierarchical Clustering',
                x = '-log10(Adj.P-val)',
                y = 'GO Term',
                alpha = 'Cluster\nMedian\nHierarchical\nhSAINT\nScore',
                size = 'FoldEnrichment',
                color = 'GO Category',
                shape = 'IsEnrichedcluster') +
  ggplot2::theme_bw() +
  ggplot2::theme(title = ggplot2::element_text(face = 'bold', size = 12),
                 axis.text = ggplot2::element_text(face = 'bold', size = 10),
                 strip.text = ggplot2::element_text(face = 'bold', size = 9), 
                 strip.background = ggplot2::element_rect(color = 'black',
                                                          fill = 'white')) +
  ggplot2::guides(#shape = ggplot2::guide_legend(override.aes = list(size = 4, shape = c(19, 17))),
                  color = ggplot2::guide_legend(override.aes = list(size = 4,
                                                                    shape = 19,
                                                                    color = c('navy',
                                                                              'darkgoldenrod2',
                                                                              'firebrick'))),
                  alpha = ggplot2::guide_legend(override.aes = list(size = 5,
                                                                    shape = 19)),
                  size = ggplot2::guide_legend(override.aes = list(shape = 19))
                  # size = 'none'
                  ) +
  ggplot2::scale_color_manual(values = c('navy',
                                         'darkgoldenrod2',
                                        'firebrick')) +
  ggplot2::scale_fill_manual(values = c('navy',
                                        'darkgoldenrod2',
                                        'firebrick'))
```

## Cluster GO (similarity) Network

```{r ClusterGOTermSimilarityNetwork, warning = F, message = F}
coords <- igraph::layout_with_fr(g)
coords <- coords*200

visNetwork::visNetwork(
  nodes = data.frame(
    id    = igraph::V(g)$name,
    label = igraph::V(g)$name,
    size = scales::rescale(
      (\(x){ x[is.na(x)] <- 0; x })(igraph::V(g)$cluster_hSAINT_score_mass),
      to = c(20, 60)
    ),
    x = coords[,1],
    y = coords[,2],
    group = base::interaction(
                        {
                          x <- igraph::V(g)$cluster
                          if (length(x) == 0) rep("", igraph::vcount(g)) else x
                        },
                        {
                          x <- igraph::V(g)$subcluster
                          if (length(x) == 0) {
                            rep("", igraph::vcount(g))
                          } else {
                            ifelse(nzchar(x), x, igraph::V(g)$cluster)
                          }
                        },
                        drop = TRUE,
                        lex.order = TRUE
                      ),
    title = sapply(
      igraph::V(g),
      function(v) {
        incoming_edges <- igraph::incident(g, v, mode = "all")
        incoming_edges <- incoming_edges[
          igraph::ends(g, incoming_edges)[,2] == igraph::V(g)[v]$name
        ]
        raw_gene_strings <- igraph::E(g)[incoming_edges]$geneID
        raw_gene_strings <- unlist(raw_gene_strings, use.names = FALSE)

        genes <- raw_gene_strings |>
          as.character() |>
          (\(x) x[!is.na(x)])() |>
          strsplit("\\\\") |>
          unlist() |>
          (\(x) x[nzchar(x)])() |>
          unique()

        term <- igraph::V(g)[v]$name
        saint_value <- igraph::V(g)[v]$cluster_hSAINT_score_mass
        if (is.na(saint_value)) saint_value <- 0
        is_cluster <- nzchar(igraph::V(g)[v]$subcluster)

        if (is_cluster) {
          paste0(
            "<div style='background-color:#ffffff;color:#1f4e79;",
            "padding:6px 10px;border-radius:6px;",
            "border:1px solid #1f4e79;font-size:12px;'>",
            "<b>", term, "</b><br>",
            "<b>SAINT mass:</b> ", signif(saint_value, 3), "<br>",
            if (length(genes) > 0)
              paste0("<u>Proteins:</u><br>", paste(genes, collapse = "<br>"))
            else "",
            "</div>"
          )
        } else {
          paste0(
            "<div style='background-color:#ffffff;color:#1f4e79;",
            "padding:6px 10px;border-radius:6px;",
            "border:1px solid #1f4e79;font-size:12px;'>",
            "<b>", term, "</b><br>",
            if (length(genes) > 0)
              paste0("<u>Proteins:</u><br>", paste(genes, collapse = "<br>"))
            else "",
            "</div>"
          )
        }
      }
    )
  ),
  edges = data.frame(
    from  = igraph::as_data_frame(g, "edges")$from,
    to    = igraph::as_data_frame(g, "edges")$to,
    value = igraph::E(g)$weight,
    title = vapply(
      seq_along(igraph::E(g)),
      function(i) {
        paste0(
          "<div style='background-color:#ffffff;color:#1f4e79;",
          "padding:6px 10px;border-radius:6px;",
          "border:1px solid #1f4e79;font-size:12px;'>",
          "<b>p.adjust:</b> ", signif(igraph::E(g)$p.adjust[i], 3), "<br>",
          "<b>weight:</b> ", signif(igraph::E(g)$weight[i], 3),
          "</div>"
        )
      },
      FUN.VALUE = character(1)
    )
  )
) |>
  # visNetwork::visNodes(color = NULL) |>
  visNetwork::visEdges(arrows = "to") |>
  visNetwork::visNodes(
    scaling = list(min = 5, max = 40),
    shadow  = TRUE
  ) |>
  visNetwork::visLayout(randomSeed = 1) |>
  visNetwork::visPhysics(enabled = FALSE) |>
  visNetwork::visInteraction(
    dragView = TRUE,
    zoomView = TRUE
  ) |>
  visNetwork::visOptions(nodesIdSelection = TRUE)
rm(coords)
```

Clusters with `enriched` GO Terms are `blue`

Here, only the enriched sub-clusters were included:
the clusters with that did not reach significant enrichment broadly, but did 
reach significant enrichment when that cluster was broken up. These `Enriched` 
**sub-clusters** are **non-blue**

Removing proteins from the list of genes comparing enrichment against in **GO GSEA** 
having been identified as contamination in literature or in the `CRAPome` significantly reduces
and changes the `GO Terms` describing `clusters`

This strategy generally favors large clusters, since they are more likely to 
retain enrichment

  + This leaves **Cluster 1** and **Cluster 25** as the only clusters retaining 
  any useful signal to describe their clusters, supporting previous experiments 
  and hypotheses regarding `TMEM`

`TMEM184B`, itself, was in **Cluster 12**, which did *NOT* return any enriched GO
Terms, likely due to its membrane protein nature (and putative localization, 
where it may progress through the endomembrane system and collide with a wide
variety of proteins), "sticking" to any number of non-specific proteinss. This 
is common among TM protein IPMS

```{r, results='asis', echo=FALSE}
fmt <- knitr::pandoc_to()

if (identical(fmt, "html")) {
  cat('## Heatmaps {.tabset .tabset-pills .tabset-fade}\n')
} else if (identical(fmt, "markdown")) {
  cat('<details><summary><strong>Heatmaps</strong></summary>\n')
} else {
  cat('## Heatmaps\n')
}
```

### Raw Counts Heatmap

```{r RawZHeatmapAll, render = T}
pheatmap::pheatmap(
  IPMS_z_ordered,
  color = grDevices::colorRampPalette(c('navy',
                                        'white',
                                        'darkgoldenrod'))((200)),
  cluster_rows = hc,
  cluster_cols = F,
  show_rownames = F,
  annotation_row = data.frame(Module = clusters_opt),
  annotation_colors = annotation_colors,
  angle_col = 90,
  use_raster = F,
  border_color = NA,
  main = 'Hierarchically-Clustered\nTMEM184B IPMS TSCs, Z-scored'
)
```

### Raw Counts Heatmap (TMEM Cluster)

```{r RawZHeatmapTMEMCluster, render = T, fig.height=14}
pheatmap::pheatmap(
  IPMS_z_ordered[rownames(IPMS_z_ordered) %in% 
                   
                   c(protein_cluster_df |> 
                       dplyr::full_join(global_enrichment_results_df) |> 
                       dplyr::full_join(subcluster_enrichment_results_df) |> 
                       dplyr::select(-c(GO_Category:Count)) |> 
                       dplyr::distinct() |> 
                       dplyr::filter(
                         cluster == protein_cluster_df |> 
                           dplyr::full_join(global_enrichment_results_df) |> 
                           dplyr::full_join(subcluster_enrichment_results_df) |> 
                           dplyr::select(-c(GO_Category:Count)) |> 
                           dplyr::filter(Protein == 'TMEM184B') |> 
                           dplyr::pull(cluster)
                         ) |> 
                       dplyr::pull(Protein)
                     ),
                 ],
  color = grDevices::colorRampPalette(c('navy',
                                        'white',
                                        'darkgoldenrod'))((200)),
  cluster_rows = T,
  cluster_cols = F,
  show_rownames = T,
  fontsize_row = 5,
  annotation_row = data.frame(
    
    cluster12_Subcluster = clusters_opt[
      names(clusters_opt) %in% 
        
        c(protein_cluster_df |> 
                       dplyr::full_join(global_enrichment_results_df) |> 
                       dplyr::full_join(subcluster_enrichment_results_df) |> 
                       dplyr::select(-c(GO_Category:Count)) |> 
                       dplyr::distinct() |> 
                       dplyr::filter(
                         cluster == protein_cluster_df |> 
                           dplyr::full_join(global_enrichment_results_df) |> 
                           dplyr::full_join(subcluster_enrichment_results_df) |> 
                           dplyr::select(-c(GO_Category:Count)) |> 
                           dplyr::filter(Protein == 'TMEM184B') |> 
                           dplyr::pull(cluster)
                         ) |> dplyr::filter(!is.na(Protein)) |> 
            dplyr::pull(Protein)
                     )
      ]
    ),
  annotation_colors = list(cluster12_Subcluster = 'darkorange2'),
  angle_col = 90,
  use_raster = F,
  border_color = NA,
  main = 'Hierarchically-Clustered TMEM184B IPMS TSCs\nZ-scored, TMEM Cluster'
)
```

`TMEM`'s cluster (**12**) is not siginficantly enriched above the proteins detected by
the mass spectrometer.

Subclusters within **cluster 12** are not enriched 

### High hSAINT Score, Low Contamination Cluster Heatmap (Signal)

```{r RawZHeatmapSAINTSignalCluster, render = T}
pheatmap::pheatmap(
  IPMS_z_ordered[rownames(IPMS_z_ordered) %in% c(protein_cluster_df |> 
                   dplyr::filter(Protein == 'TMEM184B' |
                                   grepl(cluster_class,
                                         pattern = 'High hSAINT, Low')) |> 
                   dplyr::pull(Protein)),],
  color = grDevices::colorRampPalette(c('navy',
                                        'white',
                                        'darkgoldenrod'))((200)),
  cluster_rows = T,
  cluster_cols = F,
  show_rownames = F,
  annotation_row = data.frame(
    
    cluster = clusters_opt[
      names(clusters_opt) %in% 
        
        c(protein_cluster_df |> 
            dplyr::filter(Protein == 'TMEM184B' |
                            grepl(cluster_class,
                                  pattern = 'High SAINT, Low')) |> 
            dplyr::pull(Protein)
          )
      ]
    ),
  
  annotation_colors = list(cluster = c('firebrick', 'navy')),
  angle_col = 90,
  use_raster = F,
  border_color = NA,
  cutree_rows = 3,
  main = 'Hierarchically-Clustered TMEM184B IPMS TSCs\nZ-scored, High hSAINT Score, Low Contamination Cluster + TMEM'
)
```

### High hSAINT Score, High Contamination Clusters Heatmap (Sticky/Noise)

```{r RawZHeatmapSAINTHighHighClusters, render = T}
pheatmap::pheatmap(
  IPMS_z_ordered[rownames(IPMS_z_ordered) %in% c(protein_cluster_df |> 
                   dplyr::filter(Protein == 'TMEM184B' |
                                   grepl(cluster_class,
                                         pattern = 'High hSAINT, High')) |> 
                   dplyr::pull(Protein)),],
  color = grDevices::colorRampPalette(c('navy',
                                        'white',
                                        'darkgoldenrod'))((200)),
  cluster_rows = T,
  cluster_cols = F,
  show_rownames = F,
  annotation_row = data.frame(
    
    cluster = clusters_opt[
      names(clusters_opt) %in% 
        
        c(protein_cluster_df |> 
            dplyr::filter(Protein == 'TMEM184B' |
                            grepl(cluster_class,
                                  pattern = 'High hSAINT, High')) |> 
            dplyr::pull(Protein)
          )
      ]
    ),
  
  annotation_colors = list(cluster = c('darkorange',
                                      'darkviolet',
                                      'lightpink',
                                      'skyblue')),
  angle_col = 90,
  use_raster = F,
  border_color = NA,
  cutree_rows = 2,
  main = 'Hierarchically-Clustered TMEM184B IPMS TSCs\nZ-scored, High SAINT Score, High Contamination Clusters'
)
```

### Low hSAINT Score, Low Contamination Clusters Heatmap (Low Abundance)

```{r RawZHeatmapSAINTLowLowClusters, render = T}
pheatmap::pheatmap(
  IPMS_z_ordered[rownames(IPMS_z_ordered) %in% c(protein_cluster_df |> 
                   dplyr::filter(Protein == 'TMEM184B' |
                                   grepl(cluster_class,
                                         pattern = 'Low hSAINT, Low')) |> 
                   dplyr::pull(Protein)),],
  color = grDevices::colorRampPalette(c('navy',
                                        'white',
                                        'darkgoldenrod'))((200)),
  cluster_rows = T,
  cluster_cols = F,
  show_rownames = F,
  annotation_row = data.frame(
    
    cluster = clusters_opt[
      names(clusters_opt) %in% 
        
        c(protein_cluster_df |> 
            dplyr::filter(Protein == 'TMEM184B' |
                            grepl(cluster_class,
                                  pattern = 'Low hSAINT, Low')) |> 
            dplyr::pull(Protein)
          )
      ]
    ),
  
  annotation_colors = list(cluster = annotation_colors[[1]][
    clusters_opt[
      names(clusters_opt) %in% c(protein_cluster_df |> 
                   dplyr::filter(Protein == 'TMEM184B' |
                                   grepl(cluster_class,
                                         pattern = 'Low hSAINT, Low')) |> 
                   dplyr::pull(Protein))
    ] |> as.numeric() |> unique()
  ]),
  angle_col = 90,
  use_raster = F,
  border_color = NA,
  main = 'Hierarchically-Clustered TMEM184B IPMS TSCs\nZ-scored, Low hSAINT Score, Low Contamination Cluster'
)
```

### Low hSAINT Score, High Contamination Clusters Heatmap (Background)

```{r RawZHeatmapSAINTContamClusters, render = T}
pheatmap::pheatmap(
  IPMS_z_ordered[rownames(IPMS_z_ordered) %in% c(protein_cluster_df |> 
                   dplyr::filter(Protein == 'TMEM184B' |
                                   grepl(cluster_class,
                                         pattern = 'Low hSAINT, High')) |> 
                   dplyr::pull(Protein)),],
  color = grDevices::colorRampPalette(c('navy',
                                        'white',
                                        'darkgoldenrod'))((200)),
  cluster_rows = T,
  cluster_cols = F,
  show_rownames = F,
  annotation_row = data.frame(
    
    cluster = clusters_opt[
      names(clusters_opt) %in% 
        
        c(protein_cluster_df |> 
            dplyr::filter(Protein == 'TMEM184B' |
                            grepl(cluster_class,
                                  pattern = 'Low hSAINT, High')) |> 
            dplyr::pull(Protein)
          )
      ]
    ),
  
  annotation_colors = list(cluster = annotation_colors[[1]][
    clusters_opt[
      names(clusters_opt) %in% c(protein_cluster_df |> 
                   dplyr::filter(Protein == 'TMEM184B' |
                                   grepl(cluster_class,
                                         pattern = 'Low hSAINT, High')) |> 
                   dplyr::pull(Protein))
    ] |> as.numeric() |> unique()
  ]),
  angle_col = 90,
  use_raster = F,
  border_color = NA,
  main = 'Hierarchically-Clustered TMEM184B IPMS TSCs\nZ-scored, Low hSAINT Score, High Contamination Cluster'
)
```

```{r, results='asis', echo=FALSE}
fmt <- knitr::pandoc_to()

if (identical(fmt, "html")) {
  cat('# Results Tables {.tabset .tabset-pills .tabset-fade}\n')
} else if (identical(fmt, "markdown")) {
  cat('<details><summary><strong>Results Tables</strong></summary>\n')
} else {
  cat('# Results Tables\n')
}
```

```{r restore-shiny, include=FALSE}
options(htmlwidgets.shinyMode = TRUE)
```

## Clustering Results Table by Protein (Cluster 1, first 3000 rows)

```{r Print Protein Level Results Table}
DT::datatable(
  protein_cluster_df |> 
    dplyr::full_join(global_enrichment_results_df |> 
                       dplyr::select(-c(cluster_mean_hSAINT_score:cluster_total_rank))) |> 
    dplyr::full_join(subcluster_enrichment_results_df |> 
                       dplyr::select(-c(cluster_mean_hSAINT_score:cluster_total_rank))) |> 
    dplyr::rename_with(.cols = c(dplyr::contains("rank")), 
                                 ~paste0("cluster_", .x)) |> 
    dplyr::filter(!is.na(GO_Term)) |> 
    dplyr::filter(!is.na(GO_Category)) |> 
    dplyr::distinct() |> 
    dplyr::filter(cluster == 1 &
                    grepl(protein_stat_class, pattern = 'Low Contam') &
                    protein_hSAINT_score > 0.5 &
                    protein_contam_flag == 0) |> 
    dplyr::slice_head(n = 3000)
)
```

This (complete) table exists as [hClust_protein_level_GO_GSEA.rda](../../analysis/R/hClust/hClust_protein_level_GO_GSEA.rda) in [Endo/analysis/R/hClust/](../../analysis/R/hClust)

## Clustering Results Table by Cluster

```{r Print Cluster Level Results Table}
DT::datatable(
  global_enrichment_results_df |> 
  dplyr::full_join(subcluster_enrichment_results_df) |> 
  dplyr::full_join(cluster_scores) |>
  dplyr::arrange(p.adjust, FoldEnrichment)
)
```

This table exists as [hClust_cluster_level_GO_GSEA.rda](../../analysis/R/hClust/hClust_cluster_level_GO_GSEA.rda) in [Endo/analysis/R/hClust/](../../analysis/R/hClust)

```{r, results='asis', echo=FALSE}
fmt <- knitr::pandoc_to()

if (identical(fmt, "html")) {
  cat('# Summary {.tabset .tabset-pills .tabset-fade}\n')
} else if (identical(fmt, "markdown")) {
  cat('<details><summary><strong>Summary</strong></summary>\n')
} else {
  cat('#Summary\n')
}
```

**Cluster 1** has the highest volume of `hSAINT` signal across clusters, with some 
interesting potential candidates

As noted above, it is not surprising **Cluster 12**, `TMEM`'s cluster, was not 
enriched, containing a large amount of proteins with low `hSAINT` scores, likely due
to `TMEM`'s nature as a transmembrane protein, exacerbated by its localization to 
multiple organellar compartments

A more in-depth analysis of proteins across clusters where there is greater 
probability of interacting with `TMEM` is encouraged-- collecting groups of proteins,
related by localization or function, is useful for inferring data validity; 
however, data coherence **should not negate** investigating proteins with high `hSAINT`
scores or are known/likely to interact with `TMEM` or related processes

Collectively, it appears the `hierarchical SAINT` model has done a fair job at 
delineating possible signal from noise
