---
title: "HDBSCAN"
author: "Erik Larsen"
date: "`r Sys.Date()`"
always_allow_html: true
output:
  html_document:
    code_folding: hide
    toc: true
    fig_width: 10
    fig_height: 10
  pdf_document:
    fig_width: 10
    fig_height: 10
    latex_engine: xelatex
  md_document:
    variant: gfm
---

# Overview

This vignette was developed to formally document an ad-hoc cluster analysis for 
TMEM184B immunoprecipitation tandem mass-spec (IP-MS/MS) based on experiments for 
**Dr. Martha Bhattacharya's lab** at the **University of Arizona** for the
endolysosomal paper under revision (2025-02-28) for the Journal of Cell Science

This vignette documents the `Hierarchical Density-Based Spatial Clustering of Applications with Noise (HDBSCAN)` of this IPMS data

+ The goal is to cluster protein-protein interaction candidates and assess the 
clustering

+ In the event that the clustering is uninformative, we also classify proteins by 
their `hierarchical SAINT posterior probability` (score, hereafter `hSAINT score`)--
their **signal strength** according to the hierarchical latent mixture model results


+ We do this according to the density of all `hSAINT` scores

  + The density is clearly bi-modal with separation at the extremes, thus there 
  are three categories:

    + **Low** = proteins with `hSAINT score`s within the density mode at the low 
    extreme
    
    + **High** = proteins with `hSAINT score`s within the density mode at the 
    high extreme
    
    + **Medium** = proteins with `hSAINT score`s in the "valley" between the modes
    
      + We empirically determine the "valley" using **mean shift**
    
    
+ The other dimension by which we classify proteins is **contamination**:

  + Proteins identified in literature as contamination are given a **1** (not is **0**)
  
  + Proteins identified in the CRAPome repository in more than **5%** of 
  studies are given a **1**
  
    + **Low Contamination** Proteins = **0** in both literature and CRAPome
    
    + **Medium Contamination** Proteins = **1** in either literature or CRAPome
    
    + **High Contamination** Proteins = **1** in both literature and CRAPome
    
+ We apply the same empirical binning strategy at the cluster level for 
`hSAINT score`s, since the distribution of cluster median `hSAINT` scores is also
bi-modal with a middle plateau

  + Because we lose protein-level resolution, we compute contamination rates 
  (frequency of any protein with at least one contamination flag) and apply 
  binning by tertiles, identifying an extreme outlier among a unimodal distribution
  of rates
  
  + Thus, this clustering strategy allows comparison to other clustering strategies,
  and complementation/external validity to `hSAINT scores` at the protein- and 
  cluster-level
  
    + Every protein will have an `hSAINT score`, cluster assignment, contamination rate,
    classification, and cluster classification

```{r reset-shiny-mode, include=FALSE}
options(htmlwidgets.shinyMode = FALSE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.height = 12,
  dev = c("png", "pdf"),
  fig.path = "R/HDBSCAN/"
)
```

## Attach Packages

```{r Attach Packages, message=F,warning=F}
# data/wrangling
library(tidyverse)
library(Endo)
library(DT)

  ## (hdbscan)
library(dbscan)

# GSEA/viz
library(clusterProfiler)
library(org.Hs.eg.db)
library(ReactomePA)
library(AnnotationDbi)
library(cowplot)

# progress bar
library(progressr)

# parallelization
library(future)
library(future.apply)

# network viz
library(visNetwork)

options(stringsAsFactors = FALSE)
```

## Load the Pre-Computed Data

+ Classify each protein into a matrix based on its `hSAINT` score and number of 
contamination flags (**0**, **1**, **2**)

```{r Data Load}
# Load precomputed objects from preprocessing
load("R/precompute_objects.rda")
```

### Mean Shift Function

```{r Create the Mean Shift Function to Characterize hSAINT scores}
mean_shift_valley <- function(gamma_vec) {

  # kernel density estimate on [0,1]
  d <- stats::density(gamma_vec, from = 0, to = 1)

  # tibble with density and neighbors for derivative and mode detection
  df <- tibble::tibble(
    x       = d$x,
    f       = d$y,
    f_lag   = dplyr::lag(d$y),      # f[i-1]
    f_lead  = dplyr::lead(d$y)      # f[i+1]
  )

  # discrete derivative of the density (forward difference)
  # slope[i] approximates f'(x[i])
  df <- df %>%
    dplyr::mutate(
      slope = f_lead - f
    )

  # indices where both neighbors exist (needed for local maxima)
  df <- df %>%
    dplyr::mutate(
      interior = !is.na(f_lag) & !is.na(f_lead)
    )

  # local maxima (aka modes) satisfy f > f_lag and f > f_lead
  df <- df %>%
    dplyr::mutate(
      is_mode = interior & f > f_lag & f > f_lead
    )

  # extract mode indices
  mode_idx <- which(df$is_mode)

  # assume two modes: left (noise) and right (signal)
  left_mode  <- mode_idx[1]
  right_mode <- mode_idx[length(mode_idx)]

  # mean-shift basin assignment:
  # each grid point moves uphill following the sign of the slope
  basin_target <- integer(nrow(df))

  for (i in seq_len(nrow(df))) {
    j <- i
    # move right while slope is positive
    while (j < nrow(df) && df$slope[j] > 0) {
      j <- j + 1
    }
    # move left while slope is negative
    while (j > 1 && df$slope[j - 1] < 0) {
      j <- j - 1
    }
    basin_target[i] <- j
  }

  # classify each point by which mode it's closer to
  left_dist  <- abs(basin_target - left_mode)
  right_dist <- abs(basin_target - right_mode)

  class_label <- ifelse(left_dist < right_dist, 1L, 2L)

  # valley boundary is the last index classified as left
  boundary_idx <- max(which(class_label == 1L))

  # valley location (single point)
  c <- df$x[boundary_idx]

  # expand valley asymmetrically using slope magnitude
  # threshold is relative to max slope magnitude
  eps <- 0.01 * max(abs(df$slope), na.rm = TRUE)

  left_idx <- boundary_idx
  while (left_idx > 1 && abs(df$slope[left_idx - 1]) < eps) {
    left_idx <- left_idx - 1
  }

  right_idx <- boundary_idx
  while (right_idx < nrow(df) && abs(df$slope[right_idx]) < eps) {
    right_idx <- right_idx + 1
  }

  c1 <- df$x[left_idx]
  c2 <- df$x[right_idx]

  c(c1, c2)
}
```

```{r Apply the Mean Shift}
hSAINT <- hSAINT |> 
  dplyr::mutate(
    hSAINT_class = {
      cutpoints <- mean_shift_valley(score) # find the indeces of the class modes bounding the valley between them
      dplyr::case_when(
        score < cutpoints[1] ~ 'Low hSAINT', # assign classes based on post. prob
        score <= cutpoints[2] ~ 'Medium hSAINT',
        score > cutpoints[2] ~ 'High hSAINT'
      )
    },
    contam_flag = lit_flag + CRAPome_flag,
    
    # assign contamination class based on frequency across shared studies and id in lit
    contam_class = dplyr::case_when(contam_flag == 0 ~ 
                                              'Low Contamination',
                                            contam_flag == 1 ~ 
                                              'Medium Contamination',
                                            contam_flag == 2 ~ 
                                              'High Contamination'),
    
    stat_class = paste(hSAINT_class,
                               contam_class, sep = ', ')
  ) |> # assign the orthogonal statistical model-based classes
  
  dplyr::rename(hSAINT_score = score) |> 
  dplyr::relocate(contam_flag, .after = hSAINT_score) |> 
  dplyr::select(-hSAINT_class, -contam_class) |> 
  dplyr::rename_with(.cols = c(!dplyr::contains('Protein')), ~paste0('protein_', .x))

head(hSAINT)
```

```{r Second Data Load, echo = FALSE}
rm(IPMS_mat_log)
```


```{r, results='asis', echo=FALSE}
fmt <- knitr::pandoc_to()

if (identical(fmt, "html")) {
  cat('# Principal Component Analysis {.tabset .tabset-pills .tabset-fade}\n')
} else if (identical(fmt, "markdown")) {
  cat('<details><summary><strong>Principal Component Analysis</strong></summary>\n')
} else {
  cat('# Principal Component Analysis\n')
}
```
Conduct **PCA** on the `IPMS counts dataframe` to extract the 2-D matrix required by 
`hdbscan`

```{r PCA on Counts}
pca <- stats::prcomp(IPMS_counts[,2:ncol(IPMS_counts)], scale = T)

X <- pca$x[ , 1:2, drop = FALSE]
```

```{r, results='asis', echo=FALSE}
fmt <- knitr::pandoc_to()

if (identical(fmt, "html")) {
  cat('# Clustering {.tabset .tabset-pills .tabset-fade}\n')
} else if (identical(fmt, "markdown")) {
  cat('<details><summary><strong>Clustering</strong></summary>\n')
} else {
  cat('# Clustering\n')
}
```

## Detect Clusters

+ Pass the two-dimensional object through `HDBSCAN` to obtain clustering results

+ set `minPts` (minimum number of points required to form a dense region) to 
default (**10**), which provides a stable balance between detecting 
meaningful dense regions and avoiding fragmentation

+ `HDBSCAN` will identify clusters from the 2-D object as regions of high point 
density within the embedding

```{r Scan the PCA results}
hdb <- dbscan::hdbscan(x = X, minPts = 10)
```

## Extract Clusters

```{r Extract Clusters}
cluster_labels <- hdb$cluster
```

## Tabulate Clusters

```{r Tabulate Clusters}
cluster_table <- table(cluster_labels)

cluster_labels |> 
  as.data.frame() |> 
  dplyr::mutate(cluster_labels = as.character(cluster_labels)) |> 
  dplyr::group_by(cluster_labels) |> 
  dplyr::summarize(n = dplyr::n()) |> 
  dplyr::arrange(desc(n)) |> 
  dplyr::mutate(cluster_labels = dplyr::case_when(cluster_labels == '0' ~ 'noise',
                                                  TRUE ~ cluster_labels)) |> 
  head()
```

```{r, results='asis', echo=FALSE}
fmt <- knitr::pandoc_to()

if (identical(fmt, "html")) {
  cat('# Cluster Visualization {.tabset .tabset-pills .tabset-fade}\n')
} else if (identical(fmt, "markdown")) {
  cat('<details><summary><strong>Cluster Visualization</strong></summary>\n')
} else {
  cat('# Cluster Visualization\n')
}
```

+ Plot the PCs

## PCA Plot

```{r PCA Plot}
pca_results_df <- tibble::tibble(
  PC1 = X[ , 1],
  PC2 = X[ , 2],
  cluster = cluster_labels |> 
                     as.data.frame() |> 
                     dplyr::mutate(cluster_labels = as.character(
                       cluster_labels
                     ),
                     cluster_labels = dplyr::case_when(cluster_labels == '0' ~ 
                                                         'noise',
                                                       TRUE ~ 
                                                         cluster_labels)) |> 
                     dplyr::pull()
)

cowplot::plot_grid(
  pca_results_df |> 
    ggplot2::ggplot(ggplot2::aes(x = PC1, y = PC2, group = cluster, color = cluster)) +
    ggplot2::geom_point(size = 2, alpha = 0.3) +
    ggplot2::labs(
      title = 'PCA of HDBSCAN Clusters'
      ) +
    ggplot2::theme_bw() +
    ggplot2::theme(title = ggplot2::element_text(face = 'bold', size = 12),
                   axis.title = ggplot2::element_text(face = 'bold', size = 10),
                   axis.text = ggplot2::element_text(face = 'bold', size = 9),
                   legend.position = 'bottom', 
                   legend.direction = 'horizontal',
                   legend.spacing.x = unit(0.5, 'cm'))
  ,
  pca_results_df |> 
    dplyr::filter(cluster != 'noise') |> 
    ggplot2::ggplot(ggplot2::aes(x = PC1, y = PC2, group = cluster, color = cluster)) +
    ggplot2::geom_point(size = 2, alpha = 0.3) +
    ggplot2::labs(
      title = 'PCA of HDBSCAN Clusters\nwithout "noise" cluster'
      ) +
    ggplot2::theme_bw() +
    ggplot2::theme(title = ggplot2::element_text(face = 'bold', size = 12),
                   axis.title = ggplot2::element_text(face = 'bold', size = 10),
                   axis.text = ggplot2::element_text(face = 'bold', size = 9),
                   legend.position = 'bottom', 
                   legend.direction = 'horizontal',
                   legend.spacing.x = unit(0.5, 'cm'))
    ,
  ncol = 2,
  align = 'h'
)
```

```{r, results='asis', echo=FALSE}
fmt <- knitr::pandoc_to()

if (identical(fmt, "html")) {
  cat('# Cluster Descriptions {.tabset .tabset-pills .tabset-fade}\n')
} else if (identical(fmt, "markdown")) {
  cat('<details><summary><strong>Cluster Descriptions</strong></summary>\n')
} else {
  cat('# Cluster Descriptions\n')
}
```


+ In order to assess this clustering method and to also compare/validate the `hierarchical SAINT Poisson mixture model`, 
describing clusters based on the proteins within them is useful

+ Similarly-localized proteins could be expected to have similar abundances to 
each other as well as the target (**TMEM184B**)

+ Similarly-functioning proteins could also be expected to have similar abundances 
to each other as well as the target

+ Describing these clusters attained by `HDBSCAN` via gene ontology gene set 
enrichment analysis is, therefore, useful and standard practice

+ Furthermore, we may be able to assemble a `protein-protein interaction network`
from the results

+ We outline this below-- the methods and results for acquiring these clusters, 
their constituent proteins, and how we make sense of them-- visualized as
`dotplots`, `heatmaps`, and a `network`

+ We include searchable tables printed last

## Extract Clusters

Extract gene names from the `PCA and HDBSCAN`-derived clusters and convert to a 
dataframe

```{r Extract Gene Names from Clusters}
clusters_df <- pca_results_df |> 
  dplyr::select(cluster) |> 
  dplyr::mutate(Protein = IPMS_counts |> dplyr::pull(Protein), .before = 1)

rm(list = c('X', 'pca_results_df', 'pca'))
head(clusters_df)
```

## Compute Cluster Summaries

+ Include score summaries (mean and median `hSAINT score`s, ranks, and 
contamination rates)

```{r Summarize SAINT Scores and Contamination Rates}
cluster_scores <- clusters_df |>  
  dplyr::inner_join(
    hSAINT
  ) |> 
  dplyr::arrange(desc(protein_hSAINT_score)) |> 
  dplyr::group_by(cluster) |> 
  dplyr::summarize(
    # compute summary metrics
    cluster_mean_hSAINT_score = mean(protein_hSAINT_score),
    cluster_median_hSAINT_score = stats::median(protein_hSAINT_score),
    cluster_hSAINT_score_mass = sum(protein_hSAINT_score),
    cluster_size = dplyr::n(),
    cluster_lit_flag_perc = (sum(protein_lit_flag)/cluster_size)*100,
    cluster_CRAPome_flag_perc = (sum(protein_CRAPome_flag)/cluster_size)*100,
    cluster_contam_perc = cluster_lit_flag_perc + cluster_CRAPome_flag_perc) |> 
  
  dplyr::arrange(desc(cluster_mean_hSAINT_score), cluster_contam_perc) |> 
  dplyr::mutate(cluster_mean_rank = dplyr::row_number()) |> 
  dplyr::arrange(desc(cluster_median_hSAINT_score), cluster_contam_perc) |> 
  dplyr::mutate(cluster_median_rank = dplyr::row_number()) |> 
  dplyr::arrange(desc(cluster_hSAINT_score_mass), cluster_contam_perc) |> 
  dplyr::mutate(cluster_total_rank = dplyr::row_number()) |> 
  dplyr::arrange(cluster_mean_rank) |> 
  dplyr::relocate(c(cluster_lit_flag_perc,
                    cluster_CRAPome_flag_perc), .after = cluster_total_rank) |> 
  # apply mean shift function for classification; contamination is not bi-modal
  dplyr::mutate(
    cluster_hSAINT_class = {
      cutpoints <- mean_shift_valley(cluster_median_hSAINT_score) # find the indeces of the class modes
      dplyr::case_when(
        cluster_median_hSAINT_score < cutpoints[1] ~ 'Low hSAINT', # assign classes based on post. prob
        cluster_median_hSAINT_score <= cutpoints[2] ~ 'Medium hSAINT',
        cluster_median_hSAINT_score > cutpoints[2] ~ 'High hSAINT'
      )
    },
    cluster_contam_class = dplyr::case_when(
      cluster_contam_perc <= stats::quantile(
        cluster_contam_perc, 0.60) ~ "Low Contamination",
      cluster_contam_perc <= stats::quantile(
        cluster_contam_perc, 0.95) ~ "Medium Contamination",
      TRUE ~ "High Contamination"
    ),


    
    cluster_class = paste(cluster_hSAINT_class,
                          cluster_contam_class, sep = ', ') 
  ) |> 
  dplyr::select(-cluster_hSAINT_class, -cluster_contam_class) |> 
  dplyr::relocate(cluster_class, .after = cluster)

head(cluster_scores)
```

### Cluster Distributions of hSAINT and Contamination Scores

+ Plot the distributions of metrics describing clusters, independent of `GO Terms`

```{r ClusterMetricDistributions, fig.height=8}
# extract the cluster median hSAINT score distribution class indeces
hSAINT_cuts <- mean_shift_valley(cluster_scores$cluster_median_hSAINT_score)

cowplot::plot_grid(
  cluster_scores |> 
    dplyr::group_by(cluster) |> 
    ggplot2::ggplot(ggplot2::aes(x = cluster_median_hSAINT_score)) +
    ggplot2::geom_density() +
    ggplot2::geom_rug() +
    ggplot2::theme_bw() +
    ggplot2::labs(title = 'Dist. of HDBSCAN Cluster Median hSAINT Scores',
                  x = 'Cluster Median hSAINT Score',
                  y = 'Density') +
    ggplot2::geom_vline(
      xintercept = hSAINT_cuts,
      linetype = 'dashed',
      color = 'firebrick'
    ) + ggplot2::annotate(
    "text",
    x = hSAINT_cuts[1] / 2,
    y = Inf, vjust = 1.5,
    label = "Low", color = "gray30"
  ) +
  ggplot2::annotate(
    "text",
    x = mean(hSAINT_cuts),
    y = Inf, vjust = 1.5,
    label = "Medium", color = "gray30"
  ) +
  ggplot2::annotate(
    "text",
    x = (hSAINT_cuts[2] + max(cluster_scores$cluster_median_hSAINT_score)) / 2,
    y = Inf, vjust = 1.5,
    label = "High", color = "gray30"
  ),
  
  cluster_scores |> 
    dplyr::group_by(cluster) |> 
    ggplot2::ggplot(ggplot2::aes(x = cluster_contam_perc)) +
    ggplot2::geom_density() +
    ggplot2::geom_rug() +
    ggplot2::theme_bw() +
    ggplot2::labs(title = 'Dist. of HDBSCAN Cluster Contamination Rates',
                  x = 'Cluster Contamination Rates',
                  y = 'Density') +
    ggplot2::geom_vline(
      xintercept = stats::quantile(cluster_scores |> 
                                     dplyr::pull(cluster_contam_perc),
                                   c(0.60, 0.95),
                                   na.rm = T),
      linetype = 'dashed',
      color = 'firebrick'
    )  + 
  ggplot2::annotate("text",
    x = stats::quantile(cluster_scores$cluster_contam_perc, 0.10),
    y = Inf, vjust = 1.5, label = "Low", color = "gray30"
  ) +
  ggplot2::annotate("text",
    x = stats::quantile(cluster_scores$cluster_contam_perc, 0.80),
    y = Inf, vjust = 1.5, label = "Medium", color = "gray30"
  ) +
  ggplot2::annotate("text",
    x = stats::quantile(cluster_scores$cluster_contam_perc, 0.97),
    y = Inf, vjust = 1.5, label = "High", color = "gray30"
  ),
  
  nrow = 1
)
```

+ Gather all the data together:

  + Join the cluster scores data to the cluster (per protein) dataframe and the 
`hSAINT` results for classification

```{r Join the Summary dataframe to the Protein Level dataframes}
protein_cluster_df <- clusters_df |> 
  dplyr::inner_join(
    hSAINT
  ) |> 
  dplyr::full_join(
    cluster_scores |> 
      dplyr::ungroup()
  ) |> 
  dplyr::mutate(dplyr::across(c(dplyr::where(is.double)), ~ round(.x, 3))) |> 
  dplyr::relocate(cluster_class, .after = cluster) |> 
  dplyr::arrange(desc(protein_hSAINT_score), Protein) |> 
  dplyr::group_by(cluster) |> 
  dplyr::ungroup()

head(protein_cluster_df)
```

## Cluster GO GSEA

```{r Downstream Analysis}
EntrezIDs <- AnnotationDbi::select(
  org.Hs.eg.db,
  keys = clusters_df |> dplyr::distinct(Protein) |> dplyr::pull(),
  columns = 'ENTREZID',
  keytype = 'SYMBOL'
) |> 
  dplyr::filter(!is.na(2))

EntrezMap <- stats::setNames(EntrezIDs$ENTREZID, EntrezIDs$SYMBOL)
```

+ Create enrichment functions

```{r Create Enrichment Functions}
background_genes <- EntrezIDs |> # use the EntrezIDs mapped from the modules_df
  dplyr::filter(!is.na(ENTREZID)) |> 
  dplyr::pull(ENTREZID)

get_GO <- function(genes, ontology, background_genes) {
  clusterProfiler::enrichGO(
    gene = genes,
    universe = background_genes,
    OrgDb = org.Hs.eg.db,
    keyType = 'ENTREZID',
    ont = ontology,
    pAdjustMethod = 'BH',
    qvalueCutoff = 0.05,
    readable = T
  )
}

get_kegg <- function(genes, EntrezMap) {
  entrez <- EntrezMap[genes]
  entrez <- entrez[!is.na(entrez)]
  if (length(entrez)==0) {
    return(NULL)
  }
  clusterProfiler::enrichKEGG(
    gene = entrez,
    organism = 'hsa',
    pAdjustMethod = 'BH',
    qvalueCutoff = 0.05
  )
}

get_reactome <- function(genes, EntrezMap) {
  entrez <- EntrezMap[genes]
  entrez <- entrez[!is.na(entrez)]
  if (length(entrez)==0) {
    return(NULL)
  }
  ReactomePA::enrichPathway(
    gene = entrez,
    organism = 'human',
    pAdjustMethod = 'BH',
    qvalueCutoff = 0.05,
    readable = T
  )
}
```

+ Run the functions on the dataset in parallel (multiple CPUs)

```{r Parallelize Enrichment, eval = FALSE}
future::plan(future::multisession, workers = 6)
progressr::handlers(progressr::handler_txtprogressbar())

cluster_ids <- sort(unique(clusters_df$cluster))

progressr::with_progress({
  p <- progressr::progressor(steps = length(cluster_ids))

  enrichment_results <- future.apply::future_lapply(
    cluster_ids,
    function(mod) {

      # tick the progress bar
      p(sprintf("cluster %s", mod))
      
      suppressPackageStartupMessages({
          library(clusterProfiler)
          library(DOSE)
          library(org.Hs.eg.db)
          library(dplyr) 
        })
    
        get_GO <- get_GO   # export the function so it's recognized for parallelization
        protein_cluster_df <- protein_cluster_df
        clusters_df <- clusters_df

        # proteins without any contamination except for one cluster
        genes <- clusters_df |>
          dplyr::filter(cluster == mod) |>
          dplyr::inner_join(EntrezIDs |> 
                              dplyr::rename(Protein = 1)) |> # get ENTREZID's
          dplyr::inner_join(protein_cluster_df |> 
                                dplyr::filter(protein_contam_flag == 0) |> 
                              dplyr::distinct(Protein)) |> # filter out any proteins ID'd in literature or in > 5% of CRAPome studies as contaminants
          
          dplyr::distinct(ENTREZID) |>
          dplyr::pull(ENTREZID)
      
      # all proteins within a given cluster
        background_genes <- clusters_df |> 
          dplyr::filter(cluster == mod) |> 
          dplyr::inner_join(EntrezIDs |> 
                              dplyr::rename(Protein = 1)) |> 
          dplyr::distinct(ENTREZID) |> 
          dplyr::pull()

      list(
        cluster = mod,
        go_bp  = get_GO(genes, ontology = "BP", background_genes = background_genes),
        go_cc  = get_GO(genes, ontology = "CC", background_genes = background_genes),
        go_mf  = get_GO(genes, ontology = "MF", background_genes = background_genes)
        # add KEGG/Reactome later
      )
    }
  )
})

names(enrichment_results) <- cluster_ids
rm(p)
```

+ Gather enrichment results

```{r Consolidate Enrichment Results, eval = FALSE}
global_enrichment_results_df <- lapply(enrichment_results, function(x) {
  x %>%
    purrr::map_df(., .f = as.data.frame, .id = 'GO_Category') |> 
    dplyr::rename(cluster = 2, GO_Term = Description, GO_ID = ID) |> 
    tidyr::fill(cluster, .direction = 'down') |> 
    dplyr::filter(!is.na(GO_ID)) |> 
    dplyr::mutate(GO_Category = gsub(GO_Category,
                                     pattern = 'go_',
                                     replacement = '') %>% 
                    stringr::str_to_upper(.)) |> 
    tibble::remove_rownames() |> 
    dplyr::distinct()
  }
) %>%
  purrr::map_df(., .f = as.data.frame) |>  # join to the scores and hSAINT classes
  dplyr::full_join(cluster_scores) |> 
  dplyr::filter(!is.na(GO_Category)) |> 
  dplyr::group_by(cluster) |> 
  dplyr::arrange(cluster_contam_perc, desc(cluster_median_hSAINT_score), p.adjust) |> 
  dplyr::relocate(cluster_class, .after = cluster)

head(global_enrichment_results_df)
```

```{r Load Cluster GO GSEA, eval = TRUE, echo = FALSE}
load("R/HDBSCAN/HDBSCANCluster_GO_GSEA.rda")
global_enrichment_results_df <- HDBSCANCluster_GO_GSEA
rm(HDBSCANCluster_GO_GSEA)

head(global_enrichment_results_df)
```

Removing proteins identified as contaminants by literature review or `CRAPome` 
frequency substantially alters the `GO enrichment` of clusters

  + This leaves the larger clusters as the more likely sources of enrichment
  
  + However, `HDBSCAN` will "dump" un-structured proteins into these "collection
  clusters"

**TMEM falls into the noise cluster**, likely because of its low-variance
counts across replicates and weak correlation with other protein abundances; this 
is not uncommon for a transmembrane protein with no prior for potential interactors

  + Clusters without enrichment are smaller and even small clusters with high 
  `hierarchical SAINT score mass` (cluster's `hSAINT score` sum) may not be large enough to distinguish from 
  the collection of proteins identified in the experiment

For completeness-- describing the **"un-enriched" clusters**-- "sub-cluster" 
`GO GSEA` was conducted

## SubCluster GO GSEA

Sub-cluster enrichment analysis on clusters without any enrichment (to attempt 
to describe themes of proteins within these clusters)

+ To do this, we adapt the method from the [hierarchical clustering vignette](../../analysis/HierarchicalClustering.md), where a 
dendrogram is created from the raw counts matrix

+ Then, we create a local hierarchical tree and index it by filtering *for* the 
proteins that were not in an enriched cluster

```{r SubCluster index}
"%notin%" <- Negate("%in%")

# index the clusters without significant enrichment to describe sub-clusters
tier2_clusters <- c(seq(1:48) |> as.character())[
  which(c(seq(1:48) |> as.character()) %notin% c(global_enrichment_results_df |> 
          dplyr::distinct(cluster) |> 
          dplyr::pull()) == T)] 
```

```{r SubCluster function}
# function to gather sub-clusters for each cluster that did not return enrichment results
subclusters_from_hdbscan <- function(
  counts_mat,
  clusters_df,
  n_sub_per_cluster = 3L
) {
  # extract unique HDBSCAN clusters (exclude noise = 0)
  cluster_ids <- clusters_df |> 
    dplyr::filter(cluster != 'noise') |>  
    dplyr::distinct(cluster) |> dplyr::pull()

  # list to store subcluster assignments
  out_list <- vector("list", length(cluster_ids))
  names(out_list) <- as.character(cluster_ids)

  for (k in cluster_ids) {
    # proteins in this cluster
    prots_k <- clusters_df$Protein[clusters_df$cluster == k]

    # subset the counts matrix
    sub_mat <- counts_mat[prots_k, , drop = FALSE]

    # build a local hierarchical tree
    dist_obj <- stats::dist(sub_mat)
    hc_obj   <- stats::hclust(dist_obj, method = "ward.D2")

    # number of subclusters
    K_big <- n_sub_per_cluster

    # finer cut
    sub_labels <- stats::cutree(hc_obj, k = K_big)

    # assemble tibble
    df_k <- tibble::tibble(
      Protein   = prots_k,
      cluster   = k,
      sub_label = sub_labels
    )

    # define subcluster names
    df_k <- df_k %>%
      dplyr::group_by(cluster, sub_label) %>%
      dplyr::mutate(
        subcluster = paste0("sub", dplyr::cur_group_id())
      ) %>%
      dplyr::ungroup()

    out_list[[as.character(k)]] <- df_k
  }

  # combine all clusters
  dplyr::bind_rows(out_list)
}
```

```{r SubCluster application}
# get the sub-clusters
subclusters_df <- subclusters_from_hdbscan(
  counts_mat  = IPMS_mat,
  clusters_df = dplyr::filter(clusters_df, cluster %in% tier2_clusters),
  n_sub_per_cluster = 3L
)
```

```{r Get GO GSEA on SubClusters, eval = FALSE, echo = TRUE}
future::plan(future::multisession, workers = 6)
progressr::handlers(progressr::handler_txtprogressbar())

subcluster_enrichment_df <- progressr::with_progress({

  sub_ids <- subclusters_df |> dplyr::distinct(cluster, subcluster)
  
  p <- progressr::progressor(steps = nrow(sub_ids))

  res_list <- future.apply::future_lapply(
    seq_len(nrow(sub_ids)),
    function(i) {
  
      library(clusterProfiler)
      library(dplyr)
  
      get_GO <- get_GO
      subclusters_df <- subclusters_df
      protein_cluster_df <- protein_cluster_df

      cl <- sub_ids$cluster[i]
      sc <- sub_ids$subcluster[i]

      p(sprintf("cluster %s subcluster %s", cl, sc))

      # only genes without any contamination
      genes <- subclusters_df |>
        dplyr::filter(cluster == cl, subcluster == sc) |>
        dplyr::inner_join(
          EntrezIDs |>
            dplyr::filter(!is.na(ENTREZID)) |>
            dplyr::rename(Protein = 1),
          by = "Protein"
        ) |>
        dplyr::inner_join(
          protein_cluster_df |> 
            dplyr::filter(protein_contam_flag == 0),
          by = "Protein"
        ) |>
        dplyr::distinct(ENTREZID) |>
        dplyr::pull()

      # all genes within a given cluster
      background_genes <- subclusters_df |>
        dplyr::filter(cluster == cl) |>
        dplyr::inner_join(
          EntrezIDs |>
            dplyr::filter(!is.na(ENTREZID)) |>
            dplyr::rename(Protein = 1),
          by = "Protein"
        ) |>
        dplyr::distinct(ENTREZID) |>
        dplyr::pull()

      out <- list()

      for (ont in c("BP", "CC", "MF")) {

        res <- get_GO(genes, ontology = ont, background_genes = background_genes)

        if (!is.null(res)) {
          if (nrow(res@result) > 0) {
            df <- as.data.frame(res)
            if (nrow(df) > 0) {
              df$cluster <- cl
              df$subcluster <- sc
              df$GO_Category <- ont
              out[[length(out) + 1]] <- df
            }
          }
        }
      }

      if (length(out) == 0) {
        return(NULL)
      } else {
        return(dplyr::bind_rows(out))
      }
    }
  )

  dplyr::bind_rows(res_list)
})
```


```{r Gather SubCluster GSEA data, eval = FALSE, echo = TRUE}
# wrangle the sub-cluster GO GSEA results
subcluster_enrichment_results_df <- subcluster_enrichment_df |> 
  dplyr::rename(GO_ID = ID, GO_Term = Description) |> 
  dplyr::relocate(c(cluster, subcluster, GO_Category), .before = 1) |> 
  tibble::rowid_to_column('id') |> 
  dplyr::filter(!is.na(GO_Category)) |> 
  dplyr::select(-id) |> 
  dplyr::full_join(
    cluster_scores 
  ) |>
  dplyr::filter(!is.na(GO_Category)) |> 
  dplyr::relocate(cluster_class, .after = cluster)

head(subcluster_enrichment_results_df)
```

```{r Load SubCluster GO_GSEA data, echo = F}
load("R/HDBSCAN/HDBSCANSubClustCluster_GO_GSEA.rda")
subcluster_enrichment_results_df <- HDBSCANSubClustCluster_GO_GSEA
rm(HDBSCANSubClustCluster_GO_GSEA)
subcluster_enrichment_results_df |> head()
```

```{r Find UnEnriched SubClusters}
clusters_with_subcluster_enrichment <- tier2_clusters
clusters_without_subcluster_enrichment <- generics::setdiff(clusters_with_subcluster_enrichment,
                                                          subcluster_enrichment_results_df |> 
                                                            dplyr::distinct(cluster) |> 
                                                            dplyr::pull())
clusters_with_subcluster_enrichment <- subcluster_enrichment_results_df |> 
  dplyr::distinct(cluster) |> 
  dplyr::pull()
rm(tier2_clusters)
```

## Construct the Cluster GO (similarity) Network

+ Because `HDBSCAN` produced **47 clusters** with a modest number of enriched `GO terms`,
we applied semantic similarity reduction to collapse "redundant" `GO terms` prior 
to network visualization

+ Full GO results are retained in the accompanying tables

### Gather All GO Terms

```{r Gather All GO Terms}
GO_df <- global_enrichment_results_df |> 
  dplyr::full_join(subcluster_enrichment_results_df) |> 
  dplyr::mutate(Tier = dplyr::case_when(is.na(subcluster) ~ 'Tier1',TRUE~'Tier2'),
                node_to = GO_Term,
                node_from = paste0(
                  
                  "cluster", cluster,
                  ifelse(is.na(subcluster),
                         '',
                         paste0('_Subcluster',
                                subcluster %>% 
                                  stringr::str_extract(., pattern = '\\d+')))),
                weight = -log10(p.adjust)) |> 
  dplyr::filter(weight > -log10(0.05)) |> 
  dplyr::filter(!is.na(GO_Term))

head(GO_df)
```

### Compute Semantic Similarity (Collapse)

```{r Derive a GO Term x gene Binary TabSet Matrix }
GO_gene_mat <- GO_df |> 
  dplyr::mutate(gene_set = stringr::str_split(geneID, '/')) |> 
  dplyr::select(GO_Term, gene_set) |> 
  tidyr::unnest_longer(gene_set) |> 
  stats::xtabs(~ GO_Term + gene_set, data = _)
```

```{r Compute the Jaccard Similarity Distance}
GO_jaccard_dist <- {
  A <- GO_gene_mat # local variable
  intersect_mat <- A %*% t(A) # count genes shared by every GO Term pair
  sizes <- diag(intersect_mat) # count gene size for every GO Term
  union_mat <- outer(sizes, sizes, '+') - intersect_mat # compute union of genes between GO Term pairs
  S <- intersect_mat / union_mat # compute the overlap proportion between GO Term pairs
  stats::as.dist(1 - S) # convert similarity to distance
}
```

+ Hierarchically cluster on the `GO Term jaccard distance matrix`

```{r Hierarchically Cluster Each Similarity Matrix}
GO_jaccard_clusters <- {
  hc <- stats::hclust(GO_jaccard_dist, method = 'average')
  stats::cutree(hc, h = 0.5) # assign clusters using mid-point/lit default of 0.5
}
```

+ `Jaccard clusters` are the collection of similar `GO Term` clusters per 
**HDBSCAN cluster**

+ We show all of these terms here for posterity, but will use the **5** most 
significantly enriched terms across *any* **Jaccard clusters** that are *nested within each* 
**HDBSCAN cluster** to derive and visualize a `GO Term network` for the 
`HDBSCAN clusters`

```{r Pick a representative GO Term per Jaccard Cluster}
GO_representative <- {
  df <- GO_df |> 
    dplyr::mutate(cluster_id = GO_jaccard_clusters[GO_Term]) # assign clusters
  df |> 
    dplyr::group_by(cluster_id) |> 
    dplyr::slice_min(order_by = p.adjust, n = 1, with_ties = F) |> 
    dplyr::ungroup()
}

GO_top_5_per_hdbscan <- GO_representative |> 
  dplyr::distinct(cluster_id, 
                  GO_Category, 
                  GO_Term, 
                  cluster, 
                  subcluster, Tier, FoldEnrichment, geneID, p.adjust) |> 
  dplyr::group_by(cluster) |> 
  dplyr::arrange(p.adjust) |> 
  dplyr::mutate(rank = dplyr::dense_rank(p.adjust)) |> 
  dplyr::slice_head(n = 5) |> 
  dplyr::ungroup()

GO_top_5_per_hdbscan |> head()
```

```{r Construct GO Network Object}
go_network <- GO_top_5_per_hdbscan |> 
  dplyr::full_join(cluster_scores) |> 
  dplyr::mutate(node_to = GO_Term,
                node_from = paste0(
                  
                  "cluster", cluster,
                  ifelse(is.na(subcluster),
                         '',
                         paste0('_Subcluster',
                                subcluster %>% 
                                  stringr::str_extract(., pattern = '\\d+')))),
                weight = -log10(p.adjust)) |> 
  dplyr::filter(weight > -log10(0.05)) |> 
  dplyr::filter(!is.na(GO_Term)) |> 
  dplyr::ungroup()

g <- igraph::graph_from_data_frame(
  d = go_network[, c("node_from", "node_to",
                     "weight", "geneID", "p.adjust", "FoldEnrichment")],
  directed = TRUE,
  vertices =
    dplyr::bind_rows(
      go_network |> dplyr::transmute(name = node_from, cluster_hSAINT_score_mass, subcluster),
      go_network |> dplyr::transmute(name = node_to,
                                     cluster_hSAINT_score_mass = NA_real_,
                                     subcluster = NA_character_)
    ) |>
    dplyr::distinct(name, .keep_all = TRUE)
)
```

```{r, results='asis', echo=FALSE}
fmt <- knitr::pandoc_to()

if (identical(fmt, "html")) {
  cat('# Plots {.tabset .tabset-pills .tabset-fade}\n')
} else if (identical(fmt, "markdown")) {
  cat('<details><summary><strong>Plots</strong></summary>\n')
} else {
  cat('# Plots\n')
}
```

## Cluster Dot Plot

+ Dot plot with semantic collapse of the "noise" cluster by the `Jaccard` method 
above

```{r RawClusterSubClusterDotPlotOfGOTerms, warning = F, message = F}
GO_top_5_per_hdbscan |> 
  dplyr::inner_join(
    cluster_scores |> dplyr::distinct(cluster, cluster_median_hSAINT_score)
  ) |> 
  dplyr::mutate(
    Tier = dplyr::case_when(
      is.na(subcluster) ~ paste0('Clust. ', cluster),
      TRUE ~ paste0('Clust. ', cluster,
                    ':\nsubclust. ',
                    gsub(subcluster, pattern = 'sub', replacement = ''))
      ),
    is_enriched = dplyr::case_when(is.na(subcluster) ~ 'Enriched',
                                   TRUE ~ 'Not Enriched')
    ) |> 
  dplyr::group_by(cluster) |>  
  dplyr::arrange(p.adjust) |>
  ggplot2::ggplot(
    ggplot2::aes(
      x = -log10(p.adjust),
      y = forcats::fct_rev(forcats::fct_infreq(GO_Term, cluster_median_hSAINT_score)),
      color = GO_Category,
      size = FoldEnrichment,
      alpha = cluster_median_hSAINT_score,
      shape = is_enriched,
    )
  ) +
  ggplot2::geom_vline(xintercept = -log10(0.05),
                      linetype = 'dashed',
                      color = 'firebrick',
                      size = 1,
                      alpha = 0.5) +
  ggplot2::geom_point() +
  ggplot2::facet_grid(Tier ~ ., scales = 'free_y') +
  ggplot2::labs(title = 'GO Enrichment Across\nClusters and SubClusters\nin TMEM184B IPMS\nHDBSCAN',
                x = '-log10(Adj.P-val)',
                y = 'GO Term',
                alpha = 'Cluster\nMedian\nHierarchical\nSAINT\nScore',
                size = 'FoldEnrichment',
                color = 'GO Category',
                shape = 'IsEnrichedCluster') +
  ggplot2::theme_bw() +
  ggplot2::theme(title = ggplot2::element_text(face = 'bold', size = 12),
                 axis.text = ggplot2::element_text(face = 'bold', size = 10),
                 strip.text = ggplot2::element_text(face = 'bold', size = 10), 
                 strip.background = ggplot2::element_rect(color = 'black',
                                                          fill = 'white')) +
  ggplot2::guides(#shape = ggplot2::guide_legend(override.aes = list(size = 4, shape = c(19, 17))),
                  color = ggplot2::guide_legend(override.aes = list(size = 5,
                                                                    shape = 19,
                                                                    color = c('navy',
                                                                              'darkgoldenrod2',
                                                                              'firebrick'))),
                  alpha = ggplot2::guide_legend(override.aes = list(size = 5,
                                                                    shape = 19)),
                  size = ggplot2::guide_legend(override.aes = list(shape = 19))
                  # size = 'none'
                  ) +
  ggplot2::scale_color_manual(values = c('navy',
                                         'darkgoldenrod2',
                                        'firebrick'))
```

## Cluster GO (similarity) Network

```{r ClusterGOTermSimilarityNetwork, warning = F, message = F}
coords <- igraph::layout_with_fr(g)
coords <- coords*200

visNetwork::visNetwork(
  nodes = data.frame(
    id    = igraph::V(g)$name,
    label = igraph::V(g)$name,
    size = scales::rescale(
      (\(x){ x[is.na(x)] <- 0; x })(igraph::V(g)$cluster_hSAINT_score_mass),
      to = c(20, 60)
    ),
    x = coords[,1],
    y = coords[,2],
    group = base::interaction(
                        {
                          x <- igraph::V(g)$cluster
                          if (length(x) == 0) rep("", igraph::vcount(g)) else x
                        },
                        {
                          x <- igraph::V(g)$subcluster
                          if (length(x) == 0) {
                            rep("", igraph::vcount(g))
                          } else {
                            ifelse(nzchar(x), x, igraph::V(g)$cluster)
                          }
                        },
                        drop = TRUE,
                        lex.order = TRUE
                      ),
    title = sapply(
      igraph::V(g),
      function(v) {
        incoming_edges <- igraph::incident(g, v, mode = "all")
        incoming_edges <- incoming_edges[
          igraph::ends(g, incoming_edges)[,2] == igraph::V(g)[v]$name
        ]
        raw_gene_strings <- igraph::E(g)[incoming_edges]$geneID
        raw_gene_strings <- unlist(raw_gene_strings, use.names = FALSE)

        genes <- raw_gene_strings |>
          as.character() |>
          (\(x) x[!is.na(x)])() |>
          strsplit("\\\\") |>
          unlist() |>
          (\(x) x[nzchar(x)])() |>
          unique()

        term <- igraph::V(g)[v]$name
        saint_value <- igraph::V(g)[v]$cluster_hSAINT_score_mass
        if (is.na(saint_value)) saint_value <- 0
        is_module <- nzchar(igraph::V(g)[v]$subcluster)

        if (is_module) {
          paste0(
            "<div style='background-color:#ffffff;color:#1f4e79;",
            "padding:6px 10px;border-radius:6px;",
            "border:1px solid #1f4e79;font-size:12px;'>",
            "<b>", term, "</b><br>",
            "<b>SAINT mass:</b> ", signif(saint_value, 3), "<br>",
            if (length(genes) > 0)
              paste0("<u>Proteins:</u><br>", paste(genes, collapse = "<br>"))
            else "",
            "</div>"
          )
        } else {
          paste0(
            "<div style='background-color:#ffffff;color:#1f4e79;",
            "padding:6px 10px;border-radius:6px;",
            "border:1px solid #1f4e79;font-size:12px;'>",
            "<b>", term, "</b><br>",
            if (length(genes) > 0)
              paste0("<u>Proteins:</u><br>", paste(genes, collapse = "<br>"))
            else "",
            "</div>"
          )
        }
      }
    )
  ),
  edges = data.frame(
    from  = igraph::as_data_frame(g, "edges")$from,
    to    = igraph::as_data_frame(g, "edges")$to,
    value = igraph::E(g)$weight,
    title = vapply(
      seq_along(igraph::E(g)),
      function(i) {
        paste0(
          "<div style='background-color:#ffffff;color:#1f4e79;",
          "padding:6px 10px;border-radius:6px;",
          "border:1px solid #1f4e79;font-size:12px;'>",
          "<b>p.adjust:</b> ", signif(igraph::E(g)$p.adjust[i], 3), "<br>",
          "<b>weight:</b> ", signif(igraph::E(g)$weight[i], 3),
          "</div>"
        )
      },
      FUN.VALUE = character(1)
    )
  )
) |>
  # visNetwork::visNodes(color = NULL) |>
  visNetwork::visEdges(arrows = "to") |>
  visNetwork::visNodes(
    scaling = list(min = 5, max = 40),
    shadow  = TRUE
  ) |>
  visNetwork::visLayout(randomSeed = 1) |>
  visNetwork::visPhysics(enabled = FALSE) |>
  visNetwork::visInteraction(
    dragView = TRUE,
    zoomView = TRUE
  ) |>
  visNetwork::visOptions(nodesIdSelection = TRUE)
rm(coords)
```

+ Clusters with **enriched** `GO Terms` are **blue**

```{r, results='asis', echo=FALSE}
fmt <- knitr::pandoc_to()

if (identical(fmt, "html")) {
  cat('# Results Tables {.tabset .tabset-pills .tabset-fade}\n')
} else if (identical(fmt, "markdown")) {
  cat('<details><summary><strong>Results Tables</strong></summary>\n')
} else {
  cat('# Results Tables\n')
}
```

```{r restore-shiny, include=FALSE}
options(htmlwidgets.shinyMode = TRUE)
```

## Clustering Results Table by Protein

```{r Print Protein Level Results Table}
DT::datatable(
  protein_cluster_df |> 
  dplyr::full_join(
    GO_df |> 
      dplyr::select(-c(cluster_mean_hSAINT_score:cluster_CRAPome_flag_perc, Tier:weight)) |> 
      dplyr::relocate(subcluster, .after = cluster) |> 
      dplyr::mutate(Protein = stringr::str_split(geneID, pattern = '\\/'), .before = 1) |>
      dplyr::mutate(dplyr::across(c(is.numeric),~as.character(.x))) |> 
      tidyr::pivot_longer(cols = c(2:17), names_to = 'columns', values_to = 'vals') |> 
      tidyr::unchop(cols = Protein) |> 
      tidyr::pivot_wider(id_cols = Protein, names_from = 'columns', values_from = 'vals') |> 
      tidyr::unchop(cols = c(2:17)) |> 
      dplyr::mutate(dplyr::across(c(RichFactor:qvalue, Count), ~round(as.numeric(.x), 5))) |> 
      dplyr::filter(!is.na(GO_Term))
  ) |> 
  dplyr::distinct() |> 
  dplyr::relocate(protein_stat_class, .before = protein_hSAINT_score) |> 
  dplyr::relocate(subcluster, .after = subcluster) |> 
  dplyr::relocate(c(dplyr::contains("cluster_")), .after = Count)
)
```

+ This table exists as [HDBSCAN_protein_level_GO_GSEA.rda](../../analysis/R/HDBSCAN/HDBSCAN_protein_level_GO_GSEA.rda) in [Endo/analysis/R/HDBSCAN/](../../analysis/R/HDBSCAN)

## Clustering Results Table by Cluster

```{r Print Cluster Level Results Table}
DT::datatable(
  GO_df |> 
    dplyr::full_join(cluster_scores) |> 
    dplyr::select(-c(Tier:weight)) |> 
    dplyr::relocate(subcluster, .after = cluster) |> 
    dplyr::relocate(GO_Category, .before = GO_ID) |> 
    dplyr::group_by(cluster, subcluster) |> 
    dplyr::arrange(p.adjust) |> 
    dplyr::filter(!is.na(GO_Term))
)
```

+ This table exists as [HDBSCAN_cluster_level_GO_GSEA.rda](../../analysis/R/HDBSCAN/HDBSCAN_cluster_level_GO_GSEA.rda) in [Endo/analysis/R/HDBSCAN/](../../analysis/R/HDBSCAN)

```{r, results='asis', echo=FALSE}
fmt <- knitr::pandoc_to()

if (identical(fmt, "html")) {
  cat('# Summary {.tabset .tabset-pills .tabset-fade}\n')
} else if (identical(fmt, "markdown")) {
  cat('<details><summary><strong>Summary</strong></summary>\n')
} else {
  cat('# Summary\n')
}
```


+ The largest clusters have the largest `cluster_hSAINT_score_mass` (sum of all 
`hSAINT score`s)

+ However, when sorting first by low contaminant clusters 
(`cluster_contam_perc`), **Cluster 25** has the highest volume of `hSAINT` 
signal across clusters, with some interesting potential candidates 

+ **Cluster 29** had a `cluster_hSAINT_score_mass` higher than the overall 
`cluster_hSAINT_score_mass` median score and with zero contaminants

+ Few clusters returned enriched `GO Terms`, likely due to `HDBSCAN`'s tendency to 
create many smaller-sized clusters

  + When coupled with the contamination filter on cluster member proteins 
  (further limiting the number of proteins available on which to detect 
  enrichment), this reduces the vast majority of `GO Terms`


+ As noted above, it is mildly surprising **TMEM** is in the "noise" cluster,
due to its strong consistency in abundance across replicates

+ A more in-depth analysis of proteins across clusters where there is greater 
probability of interacting with **TMEM** is encouraged-- collecting groups of proteins,
related by localization or function, is useful for inferring data validity; 
however, data coherence **should not negate** investigating proteins with high `hSAINT`
scores or are known/likely to interact with `TMEM` or related processes

+ Collectively, it appears the `hierarchical SAINT` model has done a fair job at 
delineating possible signal from noise
